---
title: "Statistical Tests of Regression: Simple Linear Regression"
subtitle: "v1.0.0"
author: "Allan Omondi"
date: today
date-format: "DD MMMM YYYY"

engine: knitr

editor: visual

format:
  html:
    toc: true
    toc-depth: 4
    toc-float: true
    number-sections: true
    fig-width: 6
    fig-height: 6
    embed-resources: true
    keep-md: false
    df-print: paged
    theme: flatly
    highlight-style: tango
  docx:
    toc: true
    toc-depth: 4
    number-sections: true
    fig-width: 6
    keep-md: false
  pdf:
    toc: true
    toc-depth: 4
    number-sections: true
    fig-width: 6
    fig-height: 6
    fig-crop: false
    keep-tex: false
    pdf-engine: xelatex

execute:
  echo: fenced
  message: false
  warning: false
  fig-width: 9
  fig-height: 5

knitr:
  opts_chunk:
    comment: "#>"
---

```{r knitr_setup}
#| message: false
#| warning: false
#| include: false

# `knitr::opts_chunk$set()` is used to set the default options for code chunks
# in a Quarto document.

# The `echo = TRUE` option means that, by default, the code will be displayed in
# the output document when it is rendered. This allows readers to see the code
# that was used to generate the results, which can be helpful for understanding
# how the analysis was performed and for reproducing the results.

# `#| echo: false` can be used to override the default options for a specific
# code chunk. This would imply that the output of the code in that chunk will
# not be displayed in the document when it is rendered. This can be useful for
# hiding code that is not relevant to the reader or for keeping the output
# document clean and focused on the results.

# An alternative is to use `#| include: false` which hides both the code and
# the output of a code chunk.

knitr::opts_chunk$set(echo = TRUE)

# `normalizePath(".")` returns the absolute path of the current working
# directory. This can be useful for ensuring that the code is running in the
# correct directory and can find the files it requires.

knitr::opts_knit$set(root.dir = normalizePath("."))

# cat("Current working directory:", getwd(), "\n")
# list.files(".")
```

# Statistical Tests of Regression: Simple Linear Regression

## What is Simple Linear Regression?

**Simple linear regression** is a statistical technique used to model the
relationship between:

- **one independent variable** `X` (predictor / explanatory variable), and

- **one dependent variable** `Y` (response / outcome variable).

Its purpose is to determine whether a **linear relationship** exists between the
two variables and, if so, to estimate that relationship with a straight-line
equation.

In this lesson:

- **Independent variable** `X`: `purchase_frequency`

- **Dependent variable** `Y`: `customer_lifetime_value`

- This means we want to examine whether customers who purchase more frequently
  tend to have a higher estimated customer lifetime value.

## The Regression Equation

Simple linear regression fits a line of the form:

$Y=\beta_0 + \beta_1X + \epsilon$

Where:

- `Y` = dependent variable (it depends on the value of `X`)

- `X` = independent variable

- $\beta_0$ ​= **intercept**

  - the predicted value of `Y` when `X = 0`

- $\beta_1$ = **slope**

  - the expected change in `Y` for a one-unit increase in `X`

- $\epsilon$ = random error term

In practice, because we work with sample data, we estimate the model as:

$\hat{Y}=b_0 + b_1X + \epsilon$

where $b_0$ and $b_1$ are the estimated regression coefficients from the sample.

## Context for This Lab

In this lab, we use simple linear regression to model the relationship between:

- **Purchase Frequency** → predictor

- **Customer Lifetime Value (CLV)** → response

We investigate whether purchase frequency is a useful predictor of customer
lifetime value and interpret the fitted regression model in business terms.

The regression model helps answer the question:

> **How does purchase frequency affect customer lifetime value? or**
>
> **How much does customer lifetime value change when purchase frequency
> increases by 1?**

If the slope $b_1$ is positive, it suggests that customers who buy more often
tend to have higher lifetime value. If it is negative, it would suggest the
opposite.

## Why Use Simple Linear Regression?

Simple linear regression is useful for two main reasons:

### 1. Explanation

It helps us understand the **direction and strength** of the relationship
between two variables.

For example:

- Is the relationship positive or negative?

- Is the relationship weak or strong?

- Is the predictor statistically significant?

### 2. Prediction

It allows us to **predict** the dependent variable from the independent
variable.

For example:

- given a customer’s purchase frequency, estimate their expected customer
  lifetime value.

## Key Outputs to Interpret

When fitting a simple linear regression model, the most important outputs are:

### 1. **Intercept (** $b_0$ )

The estimated value of `Y` when `X = 0`.

### 2. **Slope (** $b_1$ )

Shows how much `Y` is expected to change for a one-unit increase in `X`.

### 3. $R^2$ (coefficient of determination)

Shows the proportion of variation in the dependent variable explained by the
model.

- $R^2 = 0$ : the model explains none of the variation

- $R^2 = 1$ : the model explains all of the variation

### 4. **p-value for the slope**

Tests whether the independent variable (predictor) has a statistically
significant linear relationship with the dependent (response) variable.

A common hypothesis test is:

- $H_0 : \beta_1 = 0$

- $H_1 : \beta_1 \neq 0$

If the p-value is small (for example, less than 0.05), we reject $H_0$ (the null
hypothesis) and conclude that the predictor is significantly associated with the
response variable.

# Install Dependencies

```{r install_dependencies}
#| echo: true
#| message: false
#| warning: false

# `installed.packages()` retrieves a matrix of all installed packages
# `[, "Package"]` extracts on the "Package" column from the matrix of all
# packages. This contains the name of the package.
# The %in% operator is used to test if the specified package is in the matrix of
# all packages
# `dependencies = TRUE` instructs R to install not only the specified package but
# also its dependencies

if (!"pacman" %in% installed.packages()[, "Package"]) {
  install.packages("pacman", dependencies = TRUE)
  library("pacman")
}

# For reading data from a CSV file
pacman::p_load("readr")
# For data visualization
pacman::p_load("ggplot2")
# For creating a correlation plot using ggplot2
pacman::p_load("ggcorrplot")
# For creating a matrix of plots using ggplot2
pacman::p_load("GGally")
# For various statistical functions
pacman::p_load("e1071")
# For visualizing missing (Not Applicable) data
pacman::p_load("naniar")
# For reshaping data required by the correlation plot
pacman::p_load("reshape2")
# For various tests for linear regression models
pacman::p_load("lmtest")
# For quantitative validation of assumptions of linear regression models
pacman::p_load("gvlma")
```

# Load the Dataset

Execute the following R script to generate the synthetic data:
`./data/synthetic_data_for_simple_linear_regression.R`

The synthetic dataset contains the estimated Customer Lifetime Value (CLV) as
the dependent variable and the customer purchase frequency as the independent
variable. The dataset is loaded as shown below.

```{r load_dataset}
#| echo: true
#| message: false
#| warning: false

clv_data <- read_csv("./data/clv_data.csv")
head(clv_data)
```

# Initial EDA

[**View the Dimensions**]{.underline}

The number of observations and variables.

```{r show_dimensions}
#| echo: true
#| message: false
#| warning: false

dim(clv_data)
```

[**View the Data Types**]{.underline}

```{r show_data_types_1}
#| echo: true
#| message: false
#| warning: false

# sapply() is designed to apply a function to a variable in a dataset
# Example: `sapply(clv_data, class)` applies the `class()` function to each
# variable in the `clv_data` dataset, returning the data type of each variable.

sapply(clv_data, class)
```

```{r show_data_types_2}
#| echo: true
#| message: false
#| warning: false

str(clv_data)
```

[**Descriptive Statistics**]{.underline}

Descriptive statistics can be used to understand your data. Understanding your
data can lead to:

- Verifying data-entry accuracy, investigating unusual cases, deleting
  observations with outliers or imputing missing data.

- **Data transformation:** Using techniques such as scaling (dividing each value
  by the standard deviation), centering (subtracting the mean from each value),
  standardization (ensuring each numeric attribute has a mean of 0 and a
  standard deviation of 1), normalization (rescaling the values to a range of
  between 0 and 1 (inclusive)), reducing skewness using Box-Cox or Yeo-Johnson
  transformations.

- **Hypothesis formulation:** Formulate a hypothesis based on the patterns you
  identify.

- **Choosing the appropriate statistical test:** You may notice properties of
  the data such as distributions or data types that suggest the use of
  parametric or non-parametric statistical tests and parametric (linear) or
  non-parametric (non-linear) algorithms.

Typical descriptive statistics include:

1.  **Measures of frequency:** count and percent

2.  **Measures of central tendency:** mean, median, and mode

3.  **Measures of distribution/dispersion/spread/scatter/variability:** minimum,
    quartiles, maximum, variance, standard deviation, coefficient of variation,
    range, interquartile range (IQR) \[includes a box and whisker plot for
    visualization\], kurtosis, skewness \[includes a histogram for
    visualization\]).

4.  **Measures of relationship:** covariance and correlation

## [**Measures of Frequency — Not Applicable**]{.underline}

This is applicable in cases where you have categorical variables, e.g., 60% of
the observations are male and 40% are female (2 categories for the gender). This
informs you about class imbalance.

## [**Measures of Central Tendency**]{.underline}

The median and the mean of each numeric variable:

```{r central_tendency}
#| echo: true
#| message: false
#| warning: false

summary(clv_data)
```

The first 5 observations (rows) in the dataset:

```{r first_five_observations}
#| echo: true
#| message: false
#| warning: false

head(clv_data, 5)
```

The last 5 observations (rows) in the dataset:

```{r last_five_observations}
#| echo: true
#| message: false
#| warning: false

tail(clv_data, 5)
```

## [**Measures of Distribution**]{.underline}

Measuring the variability in the dataset is important because the amount of
variability determines **how well you can generalize** results from the sample
to a new observation in the population.

Low variability is ideal because it means that you can better predict
information about the population based on the sample data. High variability
means that the values are less consistent, thus making it harder to make
predictions.

The syntax `dataset[rows, columns]` can be used to specify the exact rows and
columns to be considered. `dataset[, columns]` implies all rows will be
considered. For example, specifying `BostonHousing[, -4]` implies all the
columns except column number 4. This can also be stated as
`BostonHousing[, c(1,2,3,5,6,7,8,9,10,11,12,13,14)]`. This allows us to perform
calculations on only columns that are numeric, thus leaving out the columns
termed as “factors” (categorical) or those that have a string data type.

### **Variance**

```{r distribution_variance}
#| echo: true
#| message: false
#| warning: false

sapply(clv_data[,], var)
```

### **Standard Deviation**

```{r distribution_standard_deviation}
#| echo: true
#| message: false
#| warning: false

sapply(clv_data[,], sd)
```

### **Kurtosis (Pearson)**

The Kurtosis informs us of how often outliers occur in the results. There are
different formulas for calculating kurtosis. Specifying “type = 2” allows us to
use the 2nd formula which is the same kurtosis formula used in other statistical
software like SPSS and SAS. It is referred to as "Pearson's definition of
kurtosis".

In “type = 2” (used in SPSS and SAS):

1.  Kurtosis \< 3 implies a low number of outliers → platykurtic

2.  Kurtosis = 3 implies a medium number of outliers → mesokurtic

3.  Kurtosis \> 3 implies a high number of outliers → leptokurtic

High kurtosis (leptokurtic) affects models that are sensitive to outliers.
Estimates of the variance are also inflated. Low kurtosis (platykurtic) implies
a possible underestimation of real-world variability. The typical remedy
includes trimming outliers or using robust statistical methods that are less
affected by outliers.

```{r distribution_kurtosis}
#| echo: true
#| message: false
#| warning: false

sapply(clv_data[,],  e1071::kurtosis, type = 2)
```

### **Skewness**

The skewness is used to identify the asymmetry of the distribution of results.
Similar to kurtosis, there are several ways of computing the skewness.

Using “type = 2” (common in other statistical software like SPSS and SAS) can be
interpreted as:

1.  Skewness between -0.4 and 0.4 (inclusive) implies that there is no skew in
    the distribution of results; the distribution of results is symmetrical; it
    is a normal distribution; a Gaussian distribution.

2.  Skewness above 0.4 implies a positive skew; a right-skewed distribution.

3.  Skewness below -0.4 implies a negative skew; a left-skewed distribution.

Skewed data results in misleading averages and potentially biased model
coefficients. The typical remedy to skewed data involves applying data
transformations such as Box–Cox or Yeo-Johnson transformations to reduce
skewness.

```{r distribution_skewness}
#| echo: true
#| message: false
#| warning: false

sapply(clv_data[,], e1071::skewness, type = 2)
```

**NOTE:** As a data analyst, you need to confirm if the distortion in kurtosis
or skewness is a data problem or it is a real-world insight. For example, a
real-world insight could be that few customers drive most of the value. This is
as opposed to always looking it at it as a distortion that needs to be
corrected.

## [**Measures of Relationship**]{.underline}

### **Covariance**

Covariance is a statistical measure that indicates the direction of the linear
relationship between two variables. It assesses whether increases in one
variable correspond to increases or decreases in another.​

- **Positive Covariance:** When one variable increases, the other tends to
  increase as well.

- **Negative Covariance:** When one variable increases, the other tends to
  decrease.

- **Zero Covariance:** No linear relationship exists between the variables.

While covariance indicates the direction of a relationship, it does not convey
the strength or consistency of the relationship. The correlation coefficient is
used to indicate the strength of the relationship.

```{r distribution_covariance}
#| echo: true
#| message: false
#| warning: false

cov(clv_data, method = "spearman")
```

### **Correlation**

A strong correlation between variables enables us to better predict the value of
the dependent variable using the value of the independent variable. However, a
weak correlation between two variables does not help us to predict the value of
the dependent variable from the value of the independent variable. This is
useful only if there is a linear association between the variables.

We can measure the statistical significance of the correlation using Spearman's
rank correlation *rho*. This shows us if the variables are significantly
monotonically related. A monotonic relationship between two variables implies
that as one variable increases, the other variable either consistently increases
or consistently decreases. The key characteristic is the preservation of the
direction of change, though the rate of change may vary.

```{r distribution_correlation_1}
#| echo: true
#| message: false
#| warning: false

cor.test(
  clv_data$customer_lifetime_value,
  clv_data$purchase_frequency,
  method = "spearman")
```

To view the correlation of all variables

```{r distribution_correlation_2}
#| echo: true
#| message: false
#| warning: false

cor(clv_data, method = "spearman")
```

## [**Basic Visualizations**]{.underline}

### **Histogram**

```{r visualization_histogram}
#| echo: true
#| message: false
#| warning: false

# `col_index` identifies which column(s) (by position) are being plotted;
# seq_along() generates 1, 2, ..., ncol(data_frame) automatically,
# so every column in the data frame is covered without manually listing
# positions
col_index <- seq_along(clv_data)
# the right-hand `col_index` (the full vector) is evaluated once when the
# loop starts; col_index is then reassigned to one position per iteration.
for (col_index in col_index) {
  
  # skip silently, with no message/warning, if the position does not exist
  # in the data (out of range) or if the column at that position is not
  # numeric; `next` moves straight to the following value of col_index
  if (col_index > ncol(clv_data) ||
      !is.numeric(clv_data[[col_index]])) {
    next
  }
  
  col_name <- names(clv_data)[col_index]
  
  # for a histogram, the continuous variable goes on x; geom_histogram()
  # bins it and computes the count on y automatically, so y is left
  # unmapped here
  p <- ggplot2::ggplot(clv_data, aes(x = .data[[col_name]])) +
    geom_histogram(bins = 30, fill = "#4F6EA4", color = "#FFFFFF") +
    
    # reduces the default 5% padding on the x and y axes
    scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    labs(
      title = paste0("Histogram of the Variable '", col_name, "'"),
      subtitle = paste0("EDA for distribution shape and outlier check ",
                        "before statistical tests and modelling"),
      x = col_name,
      y = "Count",
      caption = paste0("Source: CLV Data | n = ",
                       nrow(clv_data), ", ",
                       "M = ",
                       sprintf("%.2f",
                               mean(clv_data[[col_name]], na.rm = TRUE)),
                       ", ",
                       "SD = ",
                       sprintf("%.2f",
                               sd(clv_data[[col_name]], na.rm = TRUE)))
    ) +
    theme_minimal(base_size = 12) +
    
    theme(
      plot.title          = element_text(face = "bold", color = "#1D1D1D"),
      axis.title          = element_text(color = "#1D1D1D"),
      axis.text           = element_text(color = "#898989"),
      axis.text.x         = element_text(angle = 30, hjust = 1),
      plot.caption         = element_text(color = "#898989"),
      # removes vertical gridlines
      panel.grid.major.x  = element_blank(),
      panel.grid.minor.x  = element_blank(),
      # horizontal gridlines are kept (at major breaks only) since they
      # help the reader judge bar heights against the y-axis scale
      # This is part of maintaining a high data-to-ink ratio in your
      # visualizations as discussed during the theory class.
      # panel.grid.major.y  = element_blank(),
      panel.grid.minor.y  = element_blank(),
      # draws only the x-axis and y-axis lines (the "L" frame)
      axis.line.x         = element_line(color = "#1D1D1D"),
      axis.line.y         = element_line(color = "#1D1D1D")
      # panel.border = element_rect(color = "#1D1D1D", fill = NA,
      #                             linewidth = 0.5)
    )
  # ggplot objects do not auto-print inside a for-loop, therefore,
  # each one must be printed explicitly otherwise it will never appear
  print(p)
}
```

### **Box and Whisker Plot**

```{r visualization_boxplot}
#| echo: true
#| message: true
#| warning: false

# `col_index` identifies which column(s) (by position) are being plotted;
# seq_along() generates 1, 2, ..., ncol(data_frame) automatically,
# so every column in the data frame is covered without manually listing
# positions
col_index <- seq_along(clv_data)

# the right-hand `col_index` (the full vector) is evaluated once when the
# loop starts; col_index is then reassigned to one position per iteration.
for (col_index in col_index) {
  
  # skip silently, with no message/warning, if the position does not exist
  # in the data (out of range) or if the column at that position is not
  # numeric; `next` moves straight to the following value of col_index
  if (col_index > ncol(clv_data) ||
      !is.numeric(clv_data[[col_index]])) {
    next
  }
  
  col_name <- names(clv_data)[col_index]
  
  p <- ggplot2::ggplot(clv_data, aes(x = "", y = .data[[col_name]])) +
    stat_boxplot(geom = "errorbar", width = 0.15, color = "#1D1D1D") +
    geom_boxplot(fill = "#4F6EA4", color = "#1D1D1D", width = 0.3) +
    
    # reduces the default 5% padding on the x and y axes
    # scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    # scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    labs(
      title = paste0("Box and Whisker Plot of the Variable '", col_name, "'"),
      subtitle = paste0("EDA for spread and outlier check before ",
                        "statistical tests and modelling"),
      x = NULL,  # no x-axis label needed
      y = col_name,
      caption = paste0("Source: CLV Data | n = ",
                       nrow(clv_data), ", ",
                       "M = ",
                       sprintf("%.2f",
                               mean(clv_data[[col_name]], na.rm = TRUE)),
                       ", ",
                       "SD = ",
                       sprintf("%.2f",
                               sd(clv_data[[col_name]], na.rm = TRUE)))
    ) +
    theme_minimal(base_size = 12) +
    
    theme(
      plot.title          = element_text(face = "bold", color = "#1D1D1D"),
      axis.title          = element_text(color = "#1D1D1D"),
      axis.text           = element_text(color = "#898989"),
      axis.text.x         = element_text(angle = 30, hjust = 1),
      plot.caption         = element_text(color = "#898989"),
      # removes vertical gridlines
      panel.grid.major.x  = element_blank(),
      panel.grid.minor.x  = element_blank(),
      # horizontal gridlines are kept (at major breaks only) since they
      # help the reader judge bar heights against the y-axis scale
      # This is part of maintaining a high data-to-ink ratio in your
      # visualizations as discussed during the theory class.
      # panel.grid.major.y  = element_blank(),
      panel.grid.minor.y  = element_blank(),
      # draws only the x-axis and y-axis lines (the "L" frame)
      axis.line.x         = element_line(color = "#1D1D1D"),
      axis.line.y         = element_line(color = "#1D1D1D")
      # panel.border = element_rect(color = "#1D1D1D", fill = NA,
      #                             linewidth = 0.5)
    ) +
    theme_minimal(base_size = 12) +
    
    theme(
      plot.title          = element_text(face = "bold", color = "#1D1D1D"),
      axis.title          = element_text(color = "#1D1D1D"),
      axis.text           = element_text(color = "#898989"),
      axis.text.x         = element_text(angle = 30, hjust = 1),
      plot.caption         = element_text(color = "#898989"),
      # removes vertical gridlines
      panel.grid.major.x  = element_blank(),
      panel.grid.minor.x  = element_blank(),
      # horizontal gridlines are kept (at major breaks only) since they
      # help the reader judge bar heights against the y-axis scale
      # This is part of maintaining a high data-to-ink ratio in your
      # visualizations as discussed during the theory class.
      # panel.grid.major.y  = element_blank(),
      # panel.grid.minor.y  = element_blank(),
      # draws only the x-axis and y-axis lines (the "L" frame)
      axis.line.x         = element_line(color = "#1D1D1D"),
      axis.line.y         = element_line(color = "#1D1D1D")
      # panel.border = element_rect(color = "#1D1D1D", fill = NA,
      #                             linewidth = 0.5)
    )
  # ggplot objects do not auto-print inside a for-loop, therefore,
  # each one must be printed explicitly otherwise it will never appear
  print(p)
}
```

### **Missing Data Plot**

```{r missing_data_plot}
#| echo: true
#| message: false
#| warning: false

# `vis_miss()` plots a matrix of observations (rows) by variables (columns),
# shading each cell by whether the value is missing or present
naniar::vis_miss(clv_data) +
  # reduces the default 5% padding on the x and y axes
  # scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
  # scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
  labs(
    title = "Missing Data Overview",
    subtitle = "EDA for missingness pattern before data imputation or data removal",
    caption = paste0("Source: CLV Data | n = ",
                     nrow(clv_data))
  ) +
  # overrides vis_miss()'s default black/grey fill; vis_miss() has no
  # built-in color argument, but since it returns a standard ggplot
  # object, the "fill scale" can be replaced afterward
  scale_fill_manual(values = c("Present" = "#4F6EA4", "Missing" = "#E03C31")) +
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    # This is part of maintaining a high data-to-ink ratio in your
    # visualizations as discussed during the theory class.
    # panel.grid.major.y  = element_blank(),
    panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    #
  )
```

### **Correlation Plot**

```{r correlation_plot}
#| echo: true
#| message: false
#| warning: false

# Identify all numeric columns in the data frame.
# This mirrors the is.numeric() guard used inside the previouys
# boxplot and histogram loops.
# Non-numeric columns carry no meaningful Pearson correlation and are
# excluded silently before any computation begins, preventing cor() from
# throwing an error.
numeric_cols <- names(clv_data)[sapply(clv_data, is.numeric)]

# Guard: cor() requires at least two numeric columns to produce a matrix.
# Fewer than two means there is nothing to correlate. We issue an informative
# message and skip the entire block rather than letting downstream code
# fail with an uninformative error.
if (length(numeric_cols) < 2) {
  message(
    "Correlation plot skipped: fewer than two numeric columns found."
  )
} else {
  
  # Subset to numeric columns only before passing to cor().
  
  # When you subset a data frame with [, singlecolumnname], R silently drops
  # the data frame structure and returns a plain vector instead.
  
  # A vector has no columns, so cor() would immediately fail.
  # drop = FALSE suppresses that behaviour, forcing R to always return a data
  # frame regardless of how many columns are selected.
  
  # pairwise.complete.obs means each pair of variables uses all rows where
  # both values are non-missing.
  corr_matrix <- cor(
    clv_data[, numeric_cols, drop = FALSE],
    use = "pairwise.complete.obs"
  )
  
  # ggcorrplot() is a ggplot2 wrapper purpose-built for correlation matrices.
  # It returns a standard ggplot2 object, so the full suite of labs(),
  # theme_minimal(), and theme() overrides can be appended with `+`.
  #
  # hc.order = TRUE  reorders variables by hierarchical clustering so that
  #                  strongly correlated pairs appear adjacent; this makes
  #                  it easier to read than the default column order.
  # type = "lower"   shows only the lower triangle of the symmetric matrix,
  #                  thus removing redundant information.
  # lab = TRUE       prints the rounded correlation coefficient inside each
  #                  tile, allowing precise reading without relying on the
  #                  colour scale alone.
  # lab_size = 3     keeps labels legible even when many variables are present.
  p <- ggcorrplot::ggcorrplot(
    corr_matrix,
    hc.order = TRUE,
    type     = "lower",
    lab      = TRUE,
    lab_size = 3,
    
    # Diverging colour palette: negative correlations are mapped to the red
    # anchor, zero (no linear relationship) to neutral white, and positive
    # correlations to blue.
    # This preserves colour-semantic meaning (direction matters for
    # correlation in a way it does not for boxplot fills) while keeping the
    # positive anchor consistent with the house style.
    colors = c("#E03C31", "#FFFFFF", "#4F6EA4")  # negative → neutral → positive
  ) +
    
    # reduces the default 5% padding on the x and y axes
    # scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    # scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    
    labs(
      title    = "Correlation Heat-map",
      subtitle = paste0(
        "Pearson correlations, pairwise complete\n",
        "observations | Variables = ", length(numeric_cols)
      ),
      # The caption follows the same n / M / SD convention used in the
      # boxplots; here n refers to the number of observations in the data,
      # not the number of variable pairs.
      caption  = paste0("Source: CLV Data | n = ", nrow(clv_data))
    ) +
    theme_minimal(base_size = 12) +
    
    theme(
      plot.title          = element_text(face = "bold", color = "#1D1D1D"),
      axis.title          = element_text(color = "#1D1D1D"),
      axis.text           = element_text(color = "#898989"),
      axis.text.x         = element_text(angle = 30, hjust = 1),
      plot.caption         = element_text(color = "#898989"),
      # removes vertical gridlines
      panel.grid.major.x  = element_blank(),
      panel.grid.minor.x  = element_blank(),
      # horizontal gridlines are kept (at major breaks only) since they
      # help the reader judge bar heights against the y-axis scale
      # This is part of maintaining a high data-to-ink ratio in your
      # visualizations as discussed during the theory class.
      panel.grid.major.y  = element_blank(),
      panel.grid.minor.y  = element_blank(),
      # draws only the x-axis and y-axis lines (the "L" frame)
      # axis.line.x         = element_line(color = "#1D1D1D"),
      # axis.line.y         = element_line(color = "#1D1D1D")
      # panel.border = element_rect(color = "#1D1D1D", fill = NA,
      #                             linewidth = 0.5)
    )
  # ggplot objects do not auto-print inside a for-loop, therefore,
  # each one must be printed explicitly otherwise it will never appear
  print(p)
}
```

### **Scatter Plot**

```{r scatter_plot_1}
#| echo: true
#| message: false
#| warning: false

GGally::ggpairs(clv_data) +
  
  # reduces the default 5% padding on the x and y axes
  scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
  scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
  labs(
    title = "Pairwise Relationships between All Variables",
    subtitle = paste0("EDA for pairwise correlations and distributions\n",
                      "before statistical tests and before modelling.")
  ) +
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    # panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    # This is part of maintaining a high data-to-ink ratio in your
    # visualizations as discussed during the theory class.
    panel.grid.major.y  = element_blank(),
    # panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    # axis.line.x         = element_line(color = "#1D1D1D"),
    # axis.line.y         = element_line(color = "#1D1D1D")
    panel.border = element_rect(color = "#1D1D1D", fill = NA,
                               linewidth = 0.5)
  )
```

```{r scatter_plot_2, echo=TRUE, fig.width=6, message=FALSE, warning=FALSE}
pacman::p_load("ggplot2")
ggplot(clv_data,
       # It is possible to plot specific variables
       aes(x = purchase_frequency, y = customer_lifetime_value)) + 
  geom_point() +
  geom_smooth(method = lm) +
  
  # reduces the default 5% padding on the x and y axes
  scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
  scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
  
  labs(
    title = paste0("Relationship between Customer Lifetime Value and\n",
    "Purchase Frequency"),
    x = "Purchase Frequency",
    y = "Customer Lifetime Value"
  ) +
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    # This is part of maintaining a high data-to-ink ratio in your
    # visualizations as discussed during the theory class.
    # panel.grid.major.y  = element_blank(),
    # panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    #                             linewidth = 0.5)
  )
```

```{r scatter_plot_with_box_plot}
#| echo: true
#| message: false
#| warning: false

ggplot2::ggplot(clv_data,
                aes(x = "", y = customer_lifetime_value)) +
  stat_boxplot(geom = "errorbar", width = 0.15, color = "#1D1D1D") +
  # layering the raw data points underneath the box with geom_jitter()
  # shows both the summary statistics and the actual spread and sample size,
  # which a box alone hides
  geom_jitter(width = 0.1, alpha = 0.4, color = "#1D1D1D", size = 1.5) +
  geom_boxplot(fill = "#4F6EA4", color = "#1D1D1D", width = 0.4) +
  
  # reduces the default 5% padding on the x and y axes
  # scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
  # scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
  
  labs(
    title = "Distribution of Customer Lifetime Value",
    subtitle = "EDA for spread and outlier check before modelling",
    x = NULL,
    y = "Customer Lifetime Value",
    caption = paste0("Source: CLV Data | n = ", nrow(clv_data), ", ",
                     "M = ", sprintf("%.2f", mean(clv_data$customer_lifetime_value,
                                                   na.rm = TRUE)), ", ",
                     "SD = ", sprintf("%.2f", sd(clv_data$customer_lifetime_value,
                                                  na.rm = TRUE)))
  ) +
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    # This is part of maintaining a high data-to-ink ratio in your
    # visualizations as discussed during the theory class.
    # panel.grid.major.y  = element_blank(),
    # panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    #                             linewidth = 0.5)
  )
```

# Statistical Test

We then apply simple linear regression as a statistical test for regression.

```{r statistical_test_SLR}
#| echo: true
#| message: false
#| warning: false

slr_test <- lm(customer_lifetime_value ~ purchase_frequency, data = clv_data)
```

View the summary of the model.

```{r statistical_test_interpretation}
#| echo: true
#| message: false
#| warning: false

summary(slr_test)
```

A 95% confidence level means that the procedure used to construct the interval
will capture the true population parameter in 95% of repeated samples.

To obtain a 95% confidence interval:

```{r 95_confidence_interval, echo=TRUE, message=FALSE, warning=FALSE}
confint(slr_test, level = 0.95)
```

# Diagnostic EDA (Model Diagnostics)

Diagnostic Exploratory Data Analysis (EDA) is performed to verify that the
assumptions underlying the regression model are satisfied. Confirming that these
assumptions hold ensures that the statistical tests used in the model are valid
for the data, thus reducing the risk of drawing incorrect or misleading
conclusions in your data analysis.

## [**Test of Linearity**]{.underline}

The test of linearity is used to assess whether the relationship between the
dependent variable and the independent variable(s) is linear. This is necessary
given that linearity is one of the key assumptions of statistical tests of
regression and verifying it is crucial for ensuring the validity of the model's
estimates and predictions.

### **Residuals vs. Fitted Values Plot**

A plot of the residuals versus the fitted values enables us to test for
linearity. For the model to pass the test of linearity, there should be no
pattern in the distribution of residuals and the residuals should be randomly
placed around the 0.0 residual line.

```{r test_of_linearity}
#| echo: true
#| message: false
#| warning: false

pacman::p_load("ggfortify")
autoplot(slr_test, which = 1, smooth.colour = "#E03C31") +
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    # This is part of maintaining a high data-to-ink ratio in your
    # visualizations as discussed during the theory class.
    # panel.grid.major.y  = element_blank(),
    panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    #                             linewidth = 0.5)
  )
```

## [**Test of Independence of Errors (Autocorrelation)**]{.underline}

This test is necessary to confirm that each observation is independent of the
other. It helps to identify autocorrelation that is introduced when the data is
collected over a close period of time or when one observation is related to
another observation.

**Example 1:** Algorithm benchmarks run sequentially on the same machine, where
thermal throttling, caching, or background load could make consecutive runs
correlated.

**Example 2:** A retail store's daily revenue observations collected over
consecutive days may be correlated because a promotional campaign running today
continues to drive carry-over purchases tomorrow — meaning each day's revenue is
partly a function of the preceding day's revenue.

### Residuals vs. Observation Order Plot

The Residuals vs Observation Order Plot is used to assess whether the residuals
are independent of one another. The horizontal axis represents the order in
which observations appear in the dataset, while the vertical axis represents the
residual values.

**Assumption Satisfied**

If the residuals appear randomly scattered around zero with no obvious pattern,
trend, cycles, or clustering, the independence assumption is likely satisfied.

**Assumption Violated**

If the residuals show a systematic pattern, trend, wave-like structure, long
runs of positive or negative values, or clustering, the independence assumption
may be violated, indicating possible autocorrelation.

```{r residuals_vs_observation_order_plot}
#| echo: true
#| message: false
#| warning: false

residual_data <- data.frame(
  Observation = seq_along(residuals(slr_test)),
  Residual = residuals(slr_test)
)

ggplot(residual_data, aes(x = Observation, y = Residual)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Residuals vs Observation Order",
    x = "Observation Order",
    y = "Residual"
  ) +
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    # This is part of maintaining a high data-to-ink ratio in your
    # visualizations as discussed during the theory class.
    # panel.grid.major.y  = element_blank(),
    panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    #                             linewidth = 0.5)
  )
```

### Durbin-Watson Test

Autocorrelation leads to underestimated standard errors and inflated
t-statistics. It can also make findings appear more significant than they
actually are. The "Durbin-Watson Test" can be used as a test of independence of
errors (test of autocorrelation). A Durbin-Watson statistic close to 2 suggests
no autocorrelation, while values approaching 0 or 4 indicate positive or
negative autocorrelation, respectively.

For the Durbin-Watson test:

- The null hypothesis, H~0~, is that there is no autocorrelation (no
  autocorrelation = there is no correlation between residuals across time or
  across observations).

- The alternative hypothesis, H~a~, is that there is autocorrelation
  (autocorrelation = there is a correlation between residuals across time or
  across observations)

If the p-value of the Durbin-Watson statistic is greater than 0.05 then there is
no evidence to reject the null hypothesis that "there is no autocorrelation".

```{r test_of_independence_of_errors}
#| echo: true
#| message: false
#| warning: false

lmtest::dwtest(slr_test)
```

## [**Test of Normality of the Distribution of the Errors**]{.underline}

The test of normality of the distribution of the errors assesses whether the
errors (residuals) are approximately normally distributed, i.e., most errors are
close to zero and large errors are rare. A Q-Q plot can be used to conduct the
test of normality.

### Q-Q Plot

A Q-Q plot is a scatterplot of the quantiles of the errors against the quantiles
of a normal distribution. Quantiles are statistical values that divide a dataset
or probability distribution into equal-sized intervals. They help in
understanding how data is distributed by marking specific points that separate
the data into groups of equal size. Examples of quantiles include: quartiles (4
equal parts), deciles (10 equal parts), and percentiles (100 equal parts).

If the points show substantial departures from the straight diagonal line, then
the normality assumption may not be satisfied.

```{r test_of_normality}
#| echo: true
#| message: false
#| warning: false

autoplot(slr_test, which = 2, smooth.colour = NA) +
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    # This is part of maintaining a high data-to-ink ratio in your
    # visualizations as discussed during the theory class.
    # panel.grid.major.y  = element_blank(),
    panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    #                             linewidth = 0.5)
  )
```

### Shapiro-Wilk Test

The Shapiro–Wilk test is a statistical test used to determine whether the
residuals are approximately normally distributed.

The hypotheses are:

-  H₀: The residuals are normally distributed.

-  H₁: The residuals are not normally distributed.

The test produces a p-value which is compared against a chosen significance
level (usually α = 0.05).

**Decision Rule**

-  If p-value \> 0.05, fail to reject H₀. The residuals can be considered
  approximately normally distributed.

-  If p-value ≤ 0.05, reject H₀. The residuals are not normally distributed.

```{r Shapiro-Wilk-Test}
#| echo: true
#| #| message: false
#| #| warning: false

pacman::p_load("tidyverse") 
map_dfr(numeric_cols, function(col) {
  # The Shapiro-Wilk test works with data between
  # 3 observations and 5,000 observations.
  # Below is an example of how to sample 500
  # observations from the original dataset.
  test <- shapiro.test(sample(clv_data[[col]], 200))
  tibble(
    column = col,
    statistic = test$statistic,
    p_value = test$p.value
    )
  }
)
```

## [**Test of Homoscedasticity**]{.underline}

The spread of the variance of the error term should be approximately equal
across all levels of the independent variable. In the case of comparison tests,
the variances of the groups being compared should be approximately equal. A
scale-location plot (a.k.a. spread-location plot) can be used to conduct a test
of homoscedasticity.

### Scale-Location Plot

The x-axis shows the fitted (predicted) values from the model and the y-axis
shows the square root of the standardized residuals. The red line is added to
help visualize any patterns.

In a model with homoscedastic errors (equal variance across all predicted
values):

- Points should be randomly scattered around a horizontal line

- The smooth line should be approximately horizontal

- The vertical spread of points should be roughly equal across all fitted values

- No obvious patterns, funnels, or trends should be visible

Points forming a cone shape that widens from left to right suggests
heteroscedasticity with increasing variance for larger fitted values.

```{r test_of_homoscedasticity}
#| echo: true
#| message: false
#| warning: false

autoplot(slr_test, which = 3, smooth.colour = "#E03C31") +
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    # panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    # This is part of maintaining a high data-to-ink ratio in your
    # visualizations as discussed during the theory class.
    # panel.grid.major.y  = element_blank(),
    panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    #                             linewidth = 0.5)
  )
```

### **Breusch-Pagan Test**

The Breusch-Pagan Test can also be used in addition to the visual inspection of
a Scale-Location plot.

Formally:

- Null hypothesis (H₀): The spread of the variance of the error term is
  homoscedastic (equal variance).

- Alternative hypothesis (H₁): The spread of the variance of the error term is
  heteroscedastic (non-constant variance).

p-Value:

- p-value ≥ 0.05: Fail to reject H₀ → no evidence of heteroscedasticity → this
  is good; model passes.

- p-value \< 0.05: Reject H₀ → evidence of heteroscedasticity → this is bad;
  model fails.

```{r Breusch-PaganTest}
#| echo: true
#| message: false
#| warning: false

lmtest::bptest(slr_test)
```

## [**Quantitative Validation of Assumptions**]{.underline}

The graphical representations of the various tests of assumptions should be
accompanied by quantitative values. The `gvlma` package (Global Validation of
Linear Models Assumptions) is useful for this purpose.

```{r QuantitativeValidationofAssumptions}
#| echo: true
#| message: false
#| warning: false

gvlma_results <- gvlma::gvlma(slr_test)
summary(gvlma_results)
```

# Interpretation of the Results

We can interpret the results of the statistical test with more confidence if the
tests of assumptions are successful. The presentation of the results and its
subsequent interpretation is based on the following notes.

**t-Statistic t(d.f.):** It quantifies how many standard errors the estimated
coefficient deviates from zero. A larger t-value (e.g., \>2) indicates stronger
evidence against the null hypothesis (i.e., that the coefficient is zero). The
t-statistic has its corresponding p-value such that a p-value \< .05 implies a
statistically significant t-statistic.

**Degrees of Freedom (d.f.):** Degrees of freedom refers to the number of values
in a calculation that are free to vary. It is essentially a measure of how much
independent information is available for estimating a statistical parameter.

For example: Imagine you need to calculate the average height of 5 people, and
you know the sum of all their heights is 340 inches. If you know the heights of
4 of these people (65, 70, 68, and 72 inches), you can automatically determine
the height of the fifth person without measuring them: 340 - (65 + 70 + 68 + 72)
= 65 inches In this example, even though there are 5 people, you only have 4
degrees of freedom because once you know 4 heights and the total, the 5th height
is no longer “free to vary” – it is determined by the other values.

**F-Statistic**

**F(d.f. in numerator, d.f. in denominator):** The numerator degree of freedom
corresponds to the number of predictor variables, while the denominator degree
of freedom is derived from the total number of observations minus the number of
predictors and then minus 1 for the intercept.

The F-test in regression evaluates whether the variance explained by the model
is significantly greater than the unexplained variance (error). Think of the
F-statistic as a ratio of “signal” (useful prediction) to “noise” (unexplained
variation). The higher this ratio, the more confident you can be that your model
is capturing something real. The larger the F-Statistic, the better the model’s
performance.

Also, a low p-value of the F-statistic (any p-value \< .05 is considered low)
indicates that the overall regression model **is statistically significant**.

**Coefficient of Determination (R^2^)**

The R-squared value represents the proportion of the total variation in the
dependent variable that can be attributed to or explained by the independent
variable. An R-squared of 0.96 indicates that approximately 96% of the
variability in the dependent variable can be explained by its linear
relationship with the independent variable. An R-squared value approaching 1
signifies that the regression line closely aligns with the observed data points.

**Multiple R-squared:** Measures the proportion of variance in the dependent
variable explained by the independent variable (e.g., Multiple R^2^ = 0.6 means
60% of sales variance is explained by advertisement expenditure). The multiple
R-squared value always increases (or at least never decreases) when you add more
independent variables.

**Adjusted R-squared:** Also measures the proportion of variance in the
dependent variable explained by the independent variable, however, it introduces
a penalty based on the number of independent variables relative to the sample
size.

The difference between multiple R-squared and adjusted R-squared is negligible
in cases where there is only 1 independent variable.

**Residual Standard Error**

The residual standard error quantifies the average magnitude of the errors
(residuals), which are the discrepancies between the observed values in the
dataset and the values predicted by the regression model. It represents the
standard deviation of the data points around the regression line. For example, a
residual standard error of 7.73 indicates that, on average, the model's
predicted value of the dependent variable deviates from the actual observed
value by approximately 7.73 units.

A smaller residual standard error implies that the data points are more tightly
clustered around the regression line, indicating a more precise model.

**Confidence Interval**

A 95% confidence interval (CI) for a parameter—such as a regression
coefficient—provides a range that, under repeated sampling, would contain the
true (but unknown) population parameter 95% of the time. Analogy: Imagine
shooting arrows at a target. If you drew a circle around where 95% of your
arrows landed, that circle is like a confidence interval—it captures the region
in which your “shots” (i.e., estimates from different samples) tend to fall.

**Uncertainty quantification:** A CI communicates your estimate’s
precision—narrower intervals imply more precise estimates (often due to larger
samples or less variability), whereas wider intervals indicate greater
uncertainty about the true value.

**Academic Reporting (based on the APA 7th Edition Style)**

Below are some key considerations to note when reporting statistical analysis
using the APA style:\

1.  State the type of correlation test explicitly.

2.  Although not mandatory, the dependent variable is usually stated first
    followed by the independent variable when describing relationships, e.g.,
    “…to examine whether the sum of sales is collectively affected by
    advertising expenditures on YouTube, TikTok, and Facebook” such that sum of
    sales is the dependent variable that depends on advertising expenditures on
    YouTube, TikTok, and Facebook.

3.  Test statistic and parameters: Report the appropriate test statistic
    (*t*-Statistic, *F*-Statistic, $\chi^2$ , etc.) with the degrees of freedom
    in parentheses. The italicized statistical symbol is immediately followed by
    the degrees of freedom in parentheses without a space, e.g., *t*(498) and
    not t (498).

4.  Exact p-values: Report exact p-values, when possible (e.g., *p* = .032),
    unless they are less than .001, then report as *p* \< .001.

5.  Effect sizes: Include appropriate effect size measures (e.g., R², Kendall's
    W, etc.) to indicate practical significance.

6.  When the omnibus test is significant, report the post-hoc procedure used
    (e.g., the Nemenyi test)

7.  Standard errors: Report standard errors of estimates when relevant. The
    standard error tells you how much your estimate might vary if you were to
    repeat your study with different random samples from the same population. A
    smaller standard error indicates a more precise estimate.

8.  Confidence Intervals (CI): The confidence level should be clearly stated
    whenever you report point estimates (e.g., means, regression coefficients,
    correlations, etc.). The 95% confidence interval is the most common, and if
    another level is used (e.g., 90% CI, 99% CI), it should be explicitly
    mentioned. Confidence intervals are typically enclosed in square brackets
    \[\], with the lower and upper limits separated by a comma. For example: 95%
    CI \[-.03, .04\]. They are usually reported directly after the statistic
    they describe, often within the same sentence or in parentheses.

9.  Two decimal places: Report to two decimal places, except p-values which may
    need three or more decimal places.

10. Descriptive statistics: Report relevant means, standard deviations, and
    sample sizes, e.g., The sample size included 500 observations (M = 25.43, SD
    = 4.62).

11. Italicize statistical symbols: Use italics for statistical symbols (*t*,
    *F*, *p*, etc.) but not for Greek letters ( $\mu$, $\sigma$, $\alpha$),
    subscripts, or parenthetical information, e.g., R² = .45, *F*(2, 97) =
    15.62, *p* \< .001, The participants (*N* = 120) had an average score (*M* =
    25.43, *SD* = 4.62) on the cognitive test.

Further reading: <https://apastyle.apa.org/jars>

## Technical Academic Report (Based on APA Guidelines of Reporting Results)

A simple linear regression was conducted on data from 500 observations (N = 500)
to examine the relationship between customer lifetime value (CLV) and purchase
frequency. The results indicated that purchase frequency significantly predicted
CLV, $\beta$ = 19.54, 95% CI \[19.20, 19.87\], SE = 0.17, *t*(498) = 114.91, *p*
\< .001. The model explained 96.37% of the variance in CLV (R^2^ = .96, *F*(1,
498) = 13,200, *p* \< .001). For every unit increase in purchase frequency, CLV
increased by approximately 19.54 units. The intercept was 52.25, 95 % CI
\[50.48, 54.03\], and the residual standard error was 7.73, indicating strong
predictive accuracy.

## Non-Technical Business Report

The strength of the relationship highlights the critical importance of customer
retention. Initiatives that effectively encourage repeat purchases appear to be
a primary driver of customer lifetime value based on this analysis. This
understanding can guide the allocation of resources towards strategies that
foster customer loyalty and encourage repeat business.

# Rendering the Document

The "Render" button in RStudio (or `quarto render` in the terminal) can be used
to convert this Quarto document into either a:

1.  HTML document that can be opened using a browser

2.  Word Document

3.  PDF document

The conversion to PDF requires the installation of the following free software:

- For Windows: MiKTeX - <https://miktex.org/download>

- For MacOS: MacTeX - <https://www.tug.org/mactex/mactex-download.html>

- For Linux: TeX Live - <https://www.tug.org/texlive/quickinstall.html>

Also, you need to install the `tinytex` package. The `tinytex` package helps
RStudio to find and use MikTex, MacTeX, or TeXLive. Execute the following **in
the console section of RStudio** to install TinyTex:

`install.packages("tinytex")`

`tinytex::install_tinytex()`

If you are using MiKTeX for Windows, you should also enable the installation of
packages on-the-fly. This is found in “Settings \> General \> Package
Installation”

Lastly, ensure that `pdf-engine: xelatex` is set in the YAML front matter under
`format: pdf:`, as shown at the top of this document.

# References and Further Reading

American Psychological Association. (2025, February). *Journal Article Reporting
Standards (JARS)*. APA Style. Retrieved April 28, 2025, from
<https://apastyle.apa.org/jars>

Hodeghatta, U. R., & Nayak, U. (2023). *Practical Business Analytics Using R and
Python: Solve Business Problems Using a Data-driven Approach* (2nd ed.). Apress.
<https://link.springer.com/book/10.1007/978-1-4842-8754-5>
