---
title: "Statistical Tests of Comparison: ANOVA"
subtitle: "v1.0.0"
author: "Allan Omondi"
date: today
date-format: "DD MMMM YYYY"

engine: knitr

editor: visual

format:
  html:
    toc: true
    toc-depth: 4
    toc-float: true
    number-sections: true
    fig-width: 6
    fig-height: 6
    embed-resources: true
    keep-md: false
    df-print: paged
    theme: flatly
    highlight-style: tango
  docx:
    toc: true
    toc-depth: 4
    number-sections: true
    fig-width: 6
    keep-md: false
  pdf:
    toc: true
    toc-depth: 4
    number-sections: true
    fig-width: 6
    fig-height: 6
    fig-crop: false
    keep-tex: false
    pdf-engine: xelatex

execute:
  echo: fenced
  message: false
  warning: false
  fig-width: 9
  fig-height: 5

knitr:
  opts_chunk:
    comment: "#>"
---

```{r knitr_setup}
#| include: false
#| message: false
#| warning: false

# `knitr::opts_chunk$set()` is used to set the default options for code chunks
# in a Quarto document.

# The `echo = TRUE` option means that, by default, the code will be displayed in
# the output document when it is rendered. This allows readers to see the code
# that was used to generate the results, which can be helpful for understanding
# how the analysis was performed and for reproducing the results.

# `#| echo: false` can be used to override the default options for a specific
# code chunk. This would imply that the output of the code in that chunk will
# not be displayed in the document when it is rendered. This can be useful for
# hiding code that is not relevant to the reader or for keeping the output
# document clean and focused on the results.

# An alternative is to use `#| include: false` which hides both the code and
# the output of a code chunk.

knitr::opts_chunk$set(echo = TRUE)

# `normalizePath(".")` returns the absolute path of the current working
# directory. This can be useful for ensuring that the code is running in the
# correct directory and can find the files it requires.

knitr::opts_knit$set(root.dir = normalizePath("."))

# cat("Current working directory:", getwd(), "\n")
# list.files(".")
```

```{r install_dependencies}
#| echo: true
#| message: false
#| warning: false

# `installed.packages()` retrieves a matrix of all installed packages
# `[, "Package"]` extracts on the "Package" column from the matrix of all
# packages. This contains the name of the package.
# The %in% operator is used to test if the specified package is in the matrix of
# all packages
# `dependencies = TRUE` instructs R to install not only the specified package but
# also its dependencies

if (!"pacman" %in% installed.packages()[, "Package"]) {
  install.packages("pacman", dependencies = TRUE)
  library("pacman")
}

# For reading data from a CSV file
pacman::p_load("readr")
# For data visualization
pacman::p_load("ggplot2")
# For data visualization of missing data
pacman::p_load("naniar")
# "Companion to Applied Regression" for diagnostic EDA using
pacman::p_load("car")
# For the Durbin-Watson Test (test of independence of errors [autocorrelation])
pacman::p_load("lmtest")
# For performing a test of normality for MANOVA
pacman::p_load("mvnormtest")
# For visualizing hypothesis tests in multivariate linear models
pacman::p_load("heplots")
```

# Generate the Synthetic Dataset

## 4 Algorithms, 10 independent datasets, and 5 independent runs

This simulates data collected from a Machine Learning model comparison
experiment. The research question in this simulation is:

> Is there a statistically significant difference in mean classification
> accuracy across the Machine Learning algorithms?

There are 4 algorithms, 10 independent datasets, and 5 independent runs per
algorithm–dataset pair.

Execute the following R script to generate the synthetic dataset:
`./data/synthetic-data-for-ml-model-comparison.R`

“**4 × 10 × 5 = 200 observations**” means:

- 4 algorithms

- 10 [**independent datasets**]{.underline}

- 5 independent runs per algorithm–dataset pair

A "run" means:

- Different random seed

- Different initialization

- Possibly different train/test split

Therefore, the unit of analysis is the dataset, not the run. The 5 runs are
replicates, not new experimental units.

## NOT the same as Cross-Validation with Repeats

A run set up as a 10-Fold cross-validation with 5 repeats means:

For each algorithm on one dataset:

- The dataset is split into 10 folds

- The algorithm is trained and tested 10 times

- The entire 10-fold process is repeated 5 times

That gives 50 performance estimates per algorithm per dataset. However — and
this is the critical part — those 50 values are [**NOT independent
observations**]{.underline}.

They are correlated because:

- The same data points appear in multiple folds

- Training sets overlap heavily

- Test sets are reused across repeats

This disqualifies it from being used in statistical tests of comparison that
expect independence of observations.

## Valid ways to Create Independent Datasets

1.  Naturally distinct datasets
    - Different real-world sources
    - Different domains or tasks
    - Examples: Different UCI datasets, different companies’ sales datasets,
      different medical studies
2.  Independent data-generating processes
    - Simulated datasets with independent random draws
    - Different parameterizations (noise level, class imbalance, feature
      correlations)
3.  Independent experimental units
    - Different time periods where stationarity holds
    - Different geographical regions
    - Different populations

# Load the Dataset

```{r load_dataset}
#| echo: true
#| message: false
#| warning: false

ml_performance_data <- read_csv(
  "data/ml_results.csv",
  col_types = cols(
    Dataset   = col_factor(
      levels = c("Dataset_1", "Dataset_2", "Dataset_3",
                 "Dataset_4", "Dataset_5", "Dataset_6",
                 "Dataset_7", "Dataset_8", "Dataset_9",
                 "Dataset_10")),
    Run       = col_factor(levels = c("1", "2", "3", "4", "5")),
    Algorithm = col_factor(levels = c("Logistic Regression",
                                      "Random Forest",
                                      "SVM", "Gradient Boosting")),
    Accuracy  = col_double(),
    Precision = col_double(),
    Recall    = col_double()
    )
  )

head(ml_performance_data)
```

# Initial EDA

[**View the Dimensions**]{.underline}

The number of observations and variables.

```{r show_dimensions}
#| echo: true
#| message: false
#| warning: false

dim(ml_performance_data)
```

[**View the Data Types**]{.underline}

```{r show_data_types_1}
#| echo: true
#| message: false
#| warning: false

# sapply() is designed to apply a function to a variable in a dataset
# Example: `sapply(clv_data, class)` applies the `class()` function to each
# variable in the `clv_data` dataset, returning the data type of each variable.

sapply(ml_performance_data, class)
```

```{r show_data_types_2}
#| echo: true
#| message: false
#| warning: false

str(ml_performance_data)
```

[**Descriptive Statistics**]{.underline}

Understanding your data can lead to:

- **Data cleaning:** To remove extreme outliers or impute missing data.

- **Data transformation:** To reduce skewness

- **Hypothesis formulation:** Formulate a hypothesis based on the patterns you
  identify

- **Choosing the appropriate statistical test:** You may notice properties of
  the data such as distributions or data types that suggest the use of
  parametric or non-parametric statistical tests and algorithms

Descriptive statistics can be used to understand your data. Typical descriptive
statistics include:

1.  **Measures of frequency:** count and percent

2.  **Measures of central tendency:** mean, median, and mode

3.  **Measures of distribution/dispersion/spread/scatter/variability:** minimum,
    quartiles, maximum, variance, standard deviation, coefficient of variation,
    range, interquartile range (IQR) \[includes a box and whisker plot for
    visualization\], kurtosis, skewness \[includes a histogram for
    visualization\]).

4.  **Measures of relationship:** covariance and correlation \[includes a
    correlation plot and a scatter plot for visualization\].

## [**Measures of Frequency**]{.underline}

```{r measures_of_frequency}
#| echo: true
#| message: false
#| warning: false

ml_performance_data_freq <- ml_performance_data$Algorithm
cbind(frequency = table(ml_performance_data_freq),
      percentage = prop.table(table(ml_performance_data_freq)) * 100)
```

## [**Measures of Central Tendency**]{.underline}

The median and the mean of each numeric variable:

```{r central_tendency}
#| echo: true
#| message: false
#| warning: false

summary(ml_performance_data)
```

The first 5 observations (rows) in the dataset:

```{r first_five_observations}
#| echo: true
#| message: false
#| warning: false

head(ml_performance_data, 5)
```

The last 5 observations (rows) in the dataset:

```{r last_five_observations}
#| echo: true
#| message: false
#| warning: false

tail(ml_performance_data, 5)
```

## [**Measures of Distribution**]{.underline}

Measuring the variability in the dataset is important because the amount of
variability determines **how well you can generalize** results from the sample
to a new observation in the population.

Low variability is ideal because it means that you can better predict
information about the population based on the sample data. High variability
means that the values are less consistent, thus making it harder to make
predictions.

The syntax `dataset[rows, columns]` can be used to specify the exact rows and
columns to be considered. `dataset[, columns]` implies all rows will be
considered. For example, specifying `BostonHousing[, -4]` implies all the
columns except column number 4. This can also be stated as
`BostonHousing[, c(1,2,3,5,6,7,8,9,10,11,12,13,14)]`. This allows us to perform
calculations on only columns that are numeric, thus leaving out the columns
termed as “factors” (categorical) or those that have a string data type.

### **Variance**

```{r distribution_variance}
#| echo: true
#| message: false
#| warning: false

# `sapply()` is designed to apply a function to a variable in a dataset.
# In this case, we use `sapply()` to apply the `var()` function used to
# compute the variance.
sapply(ml_performance_data[,c(4)], var)
```

### **Standard Deviation**

```{r distribution_standard_deviation}
#| echo: true
#| message: false
#| warning: false

sapply(ml_performance_data[,c(4)], sd)
```

### **Kurtosis (Pearson)**

The Kurtosis informs us of how often outliers occur in the results. There are
different formulas for calculating kurtosis. Specifying “type = 2” allows us to
use the 2^nd^ formula which is the same kurtosis formula used in other
statistical software like SPSS and SAS. It is referred to as "Pearson's
definition of kurtosis".

In “type = 2” (used in SPSS and SAS):

1.  Kurtosis \< 3 implies a low number of outliers → platykurtic

2.  Kurtosis = 3 implies a medium number of outliers → mesokurtic

3.  Kurtosis \> 3 implies a high number of outliers → leptokurtic

High kurtosis (leptokurtic) affects models that are sensitive to outliers.
Estimates of the variance are also inflated. Low kurtosis (platykurtic) implies
a possible underestimation of real-world variability. The typical remedy
includes trimming outliers or using robust statistical methods that are less
affected by outliers.

```{r distribution_kurtosis}
#| echo: true
#| message: false
#| warning: false

pacman::p_load("e1071")
sapply(ml_performance_data[,c(4)],  kurtosis, type = 2)
```

### **Skewness**

The skewness is used to identify the asymmetry of the distribution of results.
Similar to kurtosis, there are several ways of computing the skewness.

Using “type = 2” (common in other statistical software like SPSS and SAS) can be
interpreted as:

1.  Skewness between -0.4 and 0.4 (inclusive) implies that there is no skew in
    the distribution of results; the distribution of results is symmetrical; it
    is a normal distribution; a Gaussian distribution.

2.  Skewness above 0.4 implies a positive skew; a right-skewed distribution.

3.  Skewness below -0.4 implies a negative skew; a left-skewed distribution.

Skewed data results in misleading averages and potentially biased model
coefficients. The typical remedy to skewed data involves applying data
transformations such as logarithmic, square-root, or Box–Cox, etc. to reduce
skewness.

```{r distribution_skewness}
#| echo: true
#| message: false
#| warning: false

sapply(ml_performance_data[,c(4)], skewness, type = 2)
```

As a data analyst, you need to confirm if the distortion in kurtosis or skewness
is a data problem or it is a real-world insight. For example, a real-world
insight could be that few customers drive most of the value. This is as opposed
to always looking it at it as a distortion that needs to be corrected.

## [**Basic Visualizations**]{.underline}

### **Histogram**

```{r visualization_histogram}
#| echo: true
#| message: false
#| warning: false

# `col_index` identifies which column(s) (by position) are being plotted;
# seq_along() generates 1, 2, ..., ncol(data_frame) automatically,
# so every column in the data frame is covered without manually listing
# positions
col_index <- seq_along(ml_performance_data)
# the right-hand `col_index` (the full vector) is evaluated once when the
# loop starts; col_index is then reassigned to one position per iteration.
for (col_index in col_index) {
  
  # skip silently, with no message/warning, if the position does not exist
  # in the data (out of range) or if the column at that position is not
  # numeric; `next` moves straight to the following value of col_index
  if (col_index > ncol(ml_performance_data) ||
      !is.numeric(ml_performance_data[[col_index]])) {
    next
  }
  
  col_name <- names(ml_performance_data)[col_index]
  
  # for a histogram, the continuous variable goes on x; geom_histogram()
  # bins it and computes the count on y automatically, so y is left
  # unmapped here (the reverse of the boxplot version, where x was a
  # blank placeholder and y held the continuous variable)
  p <- ggplot2::ggplot(ml_performance_data, aes(x = .data[[col_name]])) +
    geom_histogram(bins = 30, fill = "#4F6EA4", color = "#FFFFFF") +
    
    # reduces the default 5% padding on the x and y axes
    scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    labs(
      title = paste0("Histogram of the Variable '", col_name, "'"),
      subtitle = paste0("EDA for distribution shape and outlier check ",
                        "before statistical tests and modelling"),
      x = col_name,
      y = "Count",
      caption = paste0("Source: ML Model Performance Data | n = ",
                       nrow(ml_performance_data), ", ",
                       "M = ",
                       sprintf("%.2f",
                               mean(ml_performance_data[[col_name]], na.rm = TRUE)),
                       ", ",
                       "sd = ",
                       sprintf("%.2f",
                               sd(ml_performance_data[[col_name]], na.rm = TRUE)))
    ) +
    theme_minimal(base_size = 12) +
    
    theme(
      plot.title          = element_text(face = "bold", color = "#1D1D1D"),
      axis.title          = element_text(color = "#1D1D1D"),
      axis.text           = element_text(color = "#898989"),
      axis.text.x         = element_text(angle = 30, hjust = 1),
      plot.caption         = element_text(color = "#898989"),
      # removes vertical gridlines
      panel.grid.major.x  = element_blank(),
      panel.grid.minor.x  = element_blank(),
      # horizontal gridlines are kept (at major breaks only) since they
      # help the reader judge bar heights against the y-axis scale
      panel.grid.minor.y  = element_blank(),
      # draws only the x-axis and y-axis lines (the "L" frame)
      axis.line.x         = element_line(color = "#1D1D1D"),
      axis.line.y         = element_line(color = "#1D1D1D")
      # panel.border = element_rect(color = "#1D1D1D", fill = NA,
      #                            linewidth = 0.5)
    )
  # ggplot objects do not auto-print inside a for-loop, therefore,
  # each one must be printed explicitly otherwise it will never appear
  print(p)
}
```

### **Box and Whisker Plot**

```{r visualization_boxplot}
#| echo: true
#| message: true
#| warning: false

# `col_index` identifies which column(s) (by position) are being plotted;
# seq_along() generates 1, 2, ..., ncol(data_frame) automatically,
# so every column in the data frame is covered without manually listing
# positions
col_index <- seq_along(ml_performance_data)

# the right-hand `col_index` (the full vector) is evaluated once when the
# loop starts; col_index is then reassigned to one position per iteration.
for (col_index in col_index) {
  
  # skip silently, with no message/warning, if the position does not exist
  # in the data (out of range) or if the column at that position is not
  # numeric; `next` moves straight to the following value of col_index
  if (col_index > ncol(ml_performance_data) ||
      !is.numeric(ml_performance_data[[col_index]])) {
    next
  }
  
  col_name <- names(ml_performance_data)[col_index]
  
  p <- ggplot2::ggplot(ml_performance_data, aes(x = "", y = .data[[col_name]])) +
    stat_boxplot(geom = "errorbar", width = 0.15, color = "#1D1D1D") +
    geom_boxplot(fill = "#4F6EA4", color = "#1D1D1D", width = 0.3) +
    
    # reduces the default 5% padding on the x and y axes
    # scale_x_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    # scale_y_continuous(expand = expansion(mult = c(0.00, 0.00))) +
    labs(
      title = paste0("Box and Whisker Plot of the Variable '", col_name, "'"),
      subtitle = paste0("EDA for spread and outlier check before ",
                        "statistical tests and modelling"),
      x = NULL,  # no x-axis label needed
      y = col_name,
      caption = paste0("Source: ML Model Performance Data | n = ",
                       nrow(ml_performance_data), ", ",
                       "M = ",
                       sprintf("%.2f",
                               mean(ml_performance_data[[col_name]], na.rm = TRUE)),
                       ", ",
                       "sd = ",
                       sprintf("%.2f",
                               sd(ml_performance_data[[col_name]], na.rm = TRUE)))
    ) +
    theme_minimal(base_size = 12) +

    theme(
      plot.title          = element_text(face = "bold", color = "#1D1D1D"),
      axis.title          = element_text(color = "#1D1D1D"),
      axis.text           = element_text(color = "#898989"),
      # axis.text.x         = element_text(angle = 30, hjust = 1),
      plot.caption         = element_text(color = "#898989"),
      # removes vertical gridlines
      panel.grid.major.x  = element_blank(),
      panel.grid.minor.x  = element_blank(),
      # horizontal gridlines are kept (at major breaks only) since they
      # help the reader judge bar heights against the y-axis scale
      panel.grid.minor.y  = element_blank(),
      # draws only the x-axis and y-axis lines (the "L" frame)
      axis.line.x         = element_line(color = "#1D1D1D"),
      axis.line.y         = element_line(color = "#1D1D1D")
      # panel.border = element_rect(color = "#1D1D1D", fill = NA,
      #                            linewidth = 0.5)
    )
  # ggplot objects do not auto-print inside a for-loop, therefore,
  # each one must be printed explicitly otherwise it will never appear
  print(p)
}
```

### **Missing Data Plot**

```{r missing_data_plot}
#| echo: true
#| message: false
#| warning: false

# `vis_miss()` plots a matrix of observations (rows) by variables (columns),
# shading each cell by whether the value is missing or present; this is
# the ggplot2-native counterpart to Amelia's missmap()
naniar::vis_miss(ml_performance_data) +
  labs(
    title = "Missing Data Overview",
    subtitle = "EDA for missingness pattern before data imputation or data removal",
    caption = paste0("Source: ML Model Performance Data | n = ",
                     nrow(ml_performance_data))
  ) +
  # overrides vis_miss()'s default black/grey fill; vis_miss() has no
  # built-in color argument, but since it returns a standard ggplot
  # object, the fill scale can be replaced afterward like any other
  scale_fill_manual(values = c("Present" = "#4F6EA4", "Missing" = "#E03C31")) +
  
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    # axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge bar heights against the y-axis scale
    panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    # axis.line.x         = element_line(color = "#1D1D1D"),
    # axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
                               # linewidth = 0.5)
  )
```

### **Scatter Plot**

```{r scatter_plot_with_linear_regression_line}
#| echo: true
#| message: false
#| warning: false

ggplot2::ggplot(ml_performance_data,
                aes(x = Algorithm, y = Accuracy)) + 
  # points use the blue-grey theme discussed in class,
  # with partial transparency (alpha) so overlapping points are visible
  # rather than solid black dots which obscure density.
  geom_point(color = "#4F6EA4", alpha = 0.6, size = 2) +
  # regression line in the near-black used for borders/axis text elsewhere,
  # with a light blue-grey confidence ribbon instead of the default grey
  geom_smooth(method = lm, color = "#1D1D1D", fill = "#A9C0DA") +
  labs(
    title = "Relationship between Algorithm and Accuracy",
    subtitle = "EDA for linear association before modelling",
    x = "Algorithm",
    y = "Accuracy",
    caption = paste0("Source: ML Model Performance Data | n = ",
                     nrow(ml_performance_data))
  ) +
  
  theme_minimal(base_size = 12) +
  
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    # axis.text.x         = element_text(angle = 30, hjust = 1),
    plot.caption        = element_text(color = "#898989"),
    # removes vertical gridlines
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    # horizontal gridlines are kept (at major breaks only) since they
    # help the reader judge point position against the y-axis scale
    panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    # linewidth = 0.5)
  )
```

`geom_smooth(method = "lm")` fits a linear regression, which requires a
continuous x-variable; `Algorithm` is categorical, therefore, this is not
useful. Below is an alternative for a case where you have categorical data.

```{r scatter_plot_with_box_plot}
#| echo: true
#| message: false
#| warning: false

ggplot2::ggplot(ml_performance_data,
                aes(x = Algorithm, y = Accuracy)) +
  stat_boxplot(geom = "errorbar", width = 0.15, color = "#1D1D1D") +
  # layering the raw data points underneath the boxes with geom_jitter()
  # shows you both the summary statistics and the actual spread/sample size
  # per category, which a box alone hides, for instance you cannot tell
  # from a box alone whether one category was tested on 5 runs and another
  # on 50.
  geom_jitter(width = 0.1, alpha = 0.4, color = "#1D1D1D", size = 1.5) +
  # one box per category
  geom_boxplot(fill = "#4F6EA4", color = "#1D1D1D", width = 0.4) +
  labs(
    title = "Distribution of Accuracy by Algorithm",
    subtitle = "EDA for comparing model performance before selection",
    x = "Algorithm",
    y = "Accuracy",
    caption = paste0("Source: ML Model Performance Data | n = ",
                     nrow(ml_performance_data))
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title          = element_text(face = "bold", color = "#1D1D1D"),
    axis.title          = element_text(color = "#1D1D1D"),
    axis.text           = element_text(color = "#898989"),
    # uncomment if algorithm names are long enough to overlap when
    # printed horizontally (e.g. "Random Forest", "Gradient Boosting")
    axis.text.x        = element_text(angle = 30, hjust = 1),
    plot.caption         = element_text(color = "#898989"),
    # vertical gridlines removed: even with categorical groups on x,
    # there is no continuous scale for the lines to reference, so they
    # add no information
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    panel.grid.minor.y  = element_blank(),
    # draws only the x-axis and y-axis lines (the "L" frame)
    axis.line.x         = element_line(color = "#1D1D1D"),
    axis.line.y         = element_line(color = "#1D1D1D")
    # panel.border = element_rect(color = "#1D1D1D", fill = NA,
    # linewidth = 0.5)
  )
```

# Statistical Test: ANOVA

## One-Way ANOVA

**Purpose:** Compares the means of three or more groups defined by a single
categorical independent variable, to test whether at least one group differs
from the others.\
**Variables:** One categorical independent variable (three or more levels); one
continuous dependent variable.\
**Output:** A single F-statistic and p-value; if significant, followed by a
post-hoc test (e.g., Tukey's HSD) to identify which specific groups differ.\
**Example:** Comparing Accuracy across four algorithms used to train ML models
(accuracy depends on the chosen algorithm).

```{r one_way_anova}
#| echo: true
#| message: false
#| warning: false

one_way_anova_result <- aov(Accuracy ~ Algorithm, data = ml_performance_data)
summary(one_way_anova_result)
```

Effect size addresses a gap that the F-test and p-value do not: those tell you
whether the groups differ by more than chance, but not how much of the outcome's
variability is actually attributable to the grouping factor. Eta squared (η²)
answers that second question. It is defined as the proportion of total variance
in the dependent variable explained by the factor.

```{r eta_squared_for_one_way_anova}
#| echo: true
#| message: false
#| warning: false

# effect size (eta squared): proportion of total variance explained by 
# Algorithm; computed directly from the Sum Sq values already in the 
# summary table, no extra package required
ss <- summary(one_way_anova_result)[[1]][["Sum Sq"]]
eta_squared_oneway <- ss[1] / sum(ss)
eta_squared_oneway
```

```{r post_hoc_tukey_for_one_way_anova}
#| echo: true
#| message: false
#| warning: false

TukeyHSD(one_way_anova_result)
```

## Two-Way ANOVA

**Purpose:** Examines the effects of two categorical independent variables on a
continuous outcome simultaneously, including whether the two variables interact,
meaning the effect of one depends on the level of the other.\
**Variables:** Two categorical independent variables; one continuous dependent
variable.\
**Output:** An F-statistic and p-value for each main effect, plus a separate
F-statistic and p-value for the interaction effect; a significant interaction is
interpreted before the main effects.\
**Example:** Comparing `Accuracy` across `Algorithm` and `Dataset`, and testing
whether an algorithm's relative performance changes depending on which dataset
it was run on.

```{r two_way_anova}
#| echo: true
#| message: false
#| warning: false

ml_performance_data$Algorithm <- as.factor(ml_performance_data$Algorithm)
ml_performance_data$Dataset   <- as.factor(ml_performance_data$Dataset)

two_way_anova_result <- aov(Accuracy ~ Algorithm * Dataset, data = ml_performance_data)
shapiro.test(residuals(two_way_anova_result))
leveneTest(Accuracy ~ Algorithm * Dataset, data = ml_performance_data)

# `Algorithm * Dataset` tests both main effects and their interaction;
# a significant interaction would mean an algorithm's relative
# performance depends on which dataset it was run on, not just on
# the algorithm itself
summary(two_way_anova_result)
```

Effect size addresses a gap that the F-test and p-value do not: those tell you
whether the groups differ by more than chance, but not how much of the outcome's
variability is actually attributable to the grouping factor. Eta squared (η²)
answers that second question. It is defined as the proportion of total variance
in the dependent variable explained by the factor.

```{r eta_squared_for_two_way_anova}
#| echo: true
#| message: false
#| warning: false

ss <- summary(two_way_anova_result)[[1]][["Sum Sq"]]
row_names <- rownames(summary(two_way_anova_result)[[1]])

eta_squared_twoway <- ss / sum(ss)
names(eta_squared_twoway) <- trimws(row_names)
eta_squared_twoway
```

```{r post_hoc_tukey_for_two_way_anova}
#| message: false
#| warning: false
#| include: false

TukeyHSD(two_way_anova_result)
```

## Multiple Analysis of Variance (MANOVA)

**Purpose:** Compares group means across two or more continuous dependent
variables at the same time, accounting for the correlation between them, rather
than testing each outcome separately.\
**Variables:** One or more categorical independent variables; two or more
continuous dependent variables, analyzed jointly.\
**Output:** A multivariate test statistic (commonly Pillai's Trace); if
significant, followed by univariate ANOVAs on each dependent variable
individually to see which specific outcome(s) drove the result.\
**Example:** Comparing Accuracy, Precision, and Recall together across
Algorithm, rather than running three separate one-way ANOVAs.

```{r manova}
#| echo: true
#| message: false
#| warning: false

# requires at least two continuous DVs; this will not run on the
# current Accuracy-only dataset and is included as the template for
# when additional metrics are added per run
manova_result <- manova(cbind(Accuracy, Precision, Recall) ~ Algorithm,
                       data = ml_performance_data)

# Pillai's trace is the recommended default test statistic, since it is
# the most robust to violations of multivariate normality and unequal
# covariance matrices among the four test statistics MANOVA offers
summary(manova_result, test = "Pillai")

# univariate follow-up ANOVAs per outcome variable, the MANOVA
# equivalent of Tukey HSD follow-up after a significant one-way ANOVA
summary.aov(manova_result)
```

Effect size addresses a gap that the F-test and p-value do not: those tell you
whether the groups differ by more than chance, but not how much of the outcome's
variability is actually attributable to the grouping factor. Eta squared (η²)
answers that second question. It is defined as the proportion of total variance
in the dependent variable explained by the factor.

```{r eta_squared_for_manova}
#| echo: true
#| message: false
#| warning: false

manova_result <- manova(cbind(Accuracy, Precision, Recall) ~ Algorithm,
                        data = ml_performance_data)

aov_accuracy <- aov(Accuracy ~ Algorithm, data = ml_performance_data)
ss <- summary(aov_accuracy)[[1]][["Sum Sq"]]
eta_squared_accuracy <- ss[1] / sum(ss)
paste0("eta_squared_accuracy = ", eta_squared_accuracy)

aov_precision <- aov(Precision ~ Algorithm, data = ml_performance_data)
ss <- summary(aov_precision)[[1]][["Sum Sq"]]
eta_squared_precision <- ss[1] / sum(ss)
paste0("eta_squared_precision = ", eta_squared_precision)

aov_recall <- aov(Recall ~ Algorithm, data = ml_performance_data)
ss <- summary(aov_recall)[[1]][["Sum Sq"]]
eta_squared_recall <- ss[1] / sum(ss)
paste0("eta_squared_recall = ", eta_squared_recall)
```

```{r post_hoc_tukey_for_manova}
#| echo: true
#| message: false
#| warning: false

TukeyHSD(aov(Accuracy ~ Algorithm, data = ml_performance_data))
TukeyHSD(aov(Precision ~ Algorithm, data = ml_performance_data))
TukeyHSD(aov(Recall ~ Algorithm, data = ml_performance_data))
```

# Diagnostic EDA (Model Diagnostic)

## [**Test of Linearity — Not Applicable**]{.underline}

Linearity only exists between a predictor and outcome in a regression model.
Group membership has no inherent order or scale, so there is no "relationship"
between predictor and outcome whose shape (linear or otherwise) could be tested.

## [**Test of Independence of Errors (Autocorrelation)**]{.underline}

When observations are collected in a sequence (for example, repeated
measurements over time, or runs executed one after another), the assumption of
independence may be violated by autocorrelation: a correlation between residuals
based on their order or proximity in the sequence. The Durbin-Watson test can be
used to detect this. This test is not generally required for ANOVA designs
without a natural ordering, since independence in those cases follows from
random assignment and study design rather than from a statistical test.

Autocorrelation is consequential because, when positive, it leads to
underestimated standard errors and inflated t-statistics, which can make
findings appear more significant than they actually are. Negative
autocorrelation has the opposite effect: standard errors become overestimated
and tests become conservative.

A Durbin-Watson statistic close to 2 suggests no autocorrelation, while values
approaching 0 indicate positive autocorrelation and values approaching 4
indicate negative autocorrelation.

For the Durbin-Watson test, as implemented by default in R's `dwtest()`:

- The null hypothesis, H~0~, is that there is no autocorrelation (ρ = 0).

- The alternative hypothesis, H~a~, is that there is positive autocorrelation (ρ
  \> 0).

By default, `dwtest()` tests only for positive autocorrelation, since this is
the form most commonly encountered in sequential data. If a test against
negative autocorrelation, or against autocorrelation in either direction, is
needed instead, the `alternative` argument can be set to `"less"` or
`"two.sided"`, respectively.

If the p-value of the Durbin-Watson statistic is greater than 0.05, there is no
evidence to reject the null hypothesis of no autocorrelation.

Example: Algorithm benchmarks run sequentially on the same machine, where
thermal throttling, caching, or background load could make consecutive runs
correlated.

```{r test_of_independence_of_errors}
#| echo: true
#| message: false
#| warning: false

lmtest::dwtest(one_way_anova_result)
lmtest::dwtest(two_way_anova_result)
lmtest::dwtest(manova_result)
```

**Interpretation:** The Durbin-Watson statistic for one-way ANOVA (DW = 2.1638),
two-way ANOVA (DW=2.1142), and multiple analysis of variance (MANOVA)
(DW=2.1559) are all close to 2, indicating little evidence of autocorrelation
among the residuals. The one-way ANOVA p-value (0.8292), two-way ANOVA p-value
(0.7862), and MANOVA p-value (0.8146) are far greater than 0.05, so there is no
evidence to reject the null hypothesis. The assumption of independent errors is
therefore reasonably supported.

## [**Test of Normality of the Distribution of the Errors**]{.underline}

The Shapiro–Wilk test assesses whether a sample of data could reasonably have
come from a normally distributed population. It enables you to answer the
question, "Is this data “normal enough” for methods that assume normality?"

It tests the following hypothesis:

> H₀: The population from which the sample is drawn is normally distributed.\
> H₁: The population from which the sample is drawn is not normally distributed.

The test compares your ordered data values to the values you would expect if the
data were perfectly normal. The more they disagree, the lower the Shapiro–Wilk
statistic (W), and the smaller the p-value.

It uses the same logic as a Q–Q plot, but expressed mathematically.

**Interpretation**

If p-value \> 0.05

You fail to reject the null hypothesis. There is no evidence that the data
deviates from a normal distribution. However, this does not mean the data is
truly normal. It simply means it is "normal enough" for the chosen statistical
test of hypothesis.

**Sample Size Matters**

- Small samples (n \< 30): Shapiro–Wilk has low power. It often fails to detect
  non-normality even when it is present.

- Large samples (n \> 50): Shapiro–Wilk gains enough power to detect very small,
  practically negligible deviations from normality. A significant result at
  large n does not necessarily mean the deviation matters for your analysis;
  ANOVA is fairly robust to mild non-normality when samples are large. Pairing
  the test with a Q-Q plot helps judge whether a detected deviation is
  practically meaningful.

```{r test_of_normality}
#| echo: true
#| message: false
#| warning: false

shapiro.test(residuals(one_way_anova_result))
shapiro.test(residuals(two_way_anova_result))
result <- MVN::mvn(data = residuals(manova_result), mvn_test = "mardia")
print(result$multivariate_normality)
```

The test of normality of the distribution of the errors assesses whether the
errors (residuals) are approximately normally distributed, i.e., most errors are
close to zero and large errors are rare. A Q-Q plot can be used to conduct the
test of normality.

A Q-Q plot is a scatterplot of the quantiles of the errors against the quantiles
of a normal distribution. Quantiles are statistical values that divide a dataset
or probability distribution into equal-sized intervals. They help in
understanding how data is distributed by marking specific points that separate
the data into groups of equal size. Examples of quantiles include: quartiles (4
equal parts), percentiles (100 equal parts), deciles (10 equal parts), etc.

If the points in the Q-Q plot fall along a straight line, then the normality
assumption is satisfied. If the points in the Q-Q plot do not fall along a
straight line, then the normality assumption is not satisfied.

```{r Q-Q_plot}
#| echo: true
#| message: false
#| warning: false

plot(one_way_anova_result, which = 2)
plot(two_way_anova_result, which = 2)

resid_matrix <- residuals(manova_result)
par(mfrow = c(1, ncol(resid_matrix)))
for (i in 1:ncol(resid_matrix)) {
  qqnorm(resid_matrix[, i], main = colnames(resid_matrix)[i])
  qqline(resid_matrix[, i])
}
par(mfrow = c(1, 1))
```

**Interpretation:** The Shapiro-Wilk statistic for one-way ANOVA (W = 0.98827, p
= 0.09886) and two-way ANOVA (W = 0.99391, p = 0.587) both yield p-values
greater than 0.05, so there is no evidence to reject the null hypothesis of
normality for either model; the assumption of normally distributed residuals is
reasonably supported in both cases.

For MANOVA, residuals were instead evaluated using Mardia's multivariate
normality test, since MANOVA's normality assumption concerns the joint
distribution of residuals across all response variables rather than each
variable individually. Mardia's skewness statistic (10.708, p = 0.381) and
kurtosis statistic (0.277, p = 0.782) both yield p-values well above 0.05, so
there is no evidence to reject the null hypothesis of multivariate normality.
The assumption of normally distributed residuals is therefore reasonably
supported for the MANOVA model as well.

`shapiro.test()` evaluates only a single variable's residuals at a time, while
MANOVA's normality assumption is inherently multivariate. The Mardia test, which
properly accounts for the joint distribution, is the appropriate diagnostic here
and indicates that the assumption holds.

## [**Test of Homoscedasticity (Homogeneity of Variance)**]{.underline}

Homoscedasticity (homogeneity of variance) means that the variability of the
dependent variable is approximately the same across all groups of the
independent variable.

**Levene's Test** is used in cases where the dependent variable is numeric and
the independent variable is categorical.

It tests the following hypothesis:

> H₀: The variance of the dependent variable is equal across all levels of the
> independent variable.\
> H₁: At least one level of the independent variable has a different variance.

If the p-value is greater than 0.05, we fail to reject the null hypothesis and
conclude that there is insufficient evidence to suggest that the group variances
differ.

The purpose of checking variances is not only to satisfy a statistical
assumption. Variance can provide important practical insights about the
consistency or stability of different groups.

A group with a high variance may indicate greater uncertainty, inconsistency,
instability, or risk compared with other groups. The interpretation depends on
the application domain.

```{r test_of_homoscedasticity_one_way_anova}
#| echo: true
#| message: false
#| warning: false

car::leveneTest(Accuracy ~ Algorithm, data = ml_performance_data)
```

Meaning of the degrees of freedom from the one-way anova:

- `3` = between-group degrees of freedom, suggesting there are 4 algorithms
  (`k − 1 = 3`)

- `196` = within-group (error) degrees of freedom

This implies a total sample size of approximately:

- `N = 196 + 4 = 200`

assuming no missing values.

```{r test_of_homoscedasticity_two_way_anova}
#| echo: true
#| message: false
#| warning: false

car::leveneTest(Accuracy ~ Algorithm * Dataset, data = ml_performance_data)
```

MANOVA has multiple response variables, and the assumption it actually requires
is homogeneity of the *variance-covariance matrices* across groups, not
homogeneity of variance for each outcome individually.

```{r test_of_homoscedasticity_manova}
heplots::boxM(cbind(Accuracy, Precision, Recall) ~ Algorithm,
              data = ml_performance_data)
```

**Interpretation:**

For the one-way ANOVA, Levene's test indicated that the variances of accuracy
scores were not significantly different across the algorithms, *F*(3, 196) =
0.256, p = .857. The assumption of homogeneity of variance was therefore
satisfied.

For the two-way ANOVA, Levene's test was conducted across all combinations of
Algorithm and Dataset to assess homogeneity of variance within each cell of the
design. The result was non-significant, *F*(39, 160) = 0.908, p = .629,
indicating that the assumption of homogeneity of variance was satisfied across
the crossed design.

For the MANOVA, homogeneity of variance-covariance matrices across Algorithm
groups was assessed using Box's M test, since MANOVA's assumption concerns the
joint covariance structure across all dependent variables (Accuracy, Precision,
Recall) rather than the variance of each outcome individually. The result was
non-significant, $\chi2$ (18) = 12.360, p = .828, indicating no evidence against
the assumption of homogeneous covariance matrices across groups. The assumption
was therefore reasonably supported.

Box's M test is known to be highly sensitive to departures from multivariate
normality and tends to reject the null more readily as sample size increases; a
non-significant result here is reassuring, but a significant result would not
necessarily indicate a fatal violation, particularly with reasonably balanced
group sizes, given MANOVA's general robustness to this assumption under such
conditions.

## [**Quantitative Validation of Assumptions — Not Applicable**]{.underline}

Although `aov()` objects do generate residuals and can technically be passed to
`gvlma()`, the function's link function (linearity) component is not meaningful
here, since the indepenent variable is a categorical factor with no inherent
order or numeric scale. Quantitative validation is instead carried out using
assumption-specific tests: `shapiro.test()` for normality, `leveneTest()` for
homogeneity of variance, and, where observations are sequentially ordered,
`dwtest()` for independence.

# Interpretation of the Results

The presentation of the results and its subsequent interpretation is based on
the following notes.

**t-Statistic—t(d.f.):** In a one-sample t-test, the t-statistic measures how
far the sample mean is from the hypothesised population mean.

- A t-value close to 0 indicates that the sample mean is very close to the
  hypothesised value.

- A larger absolute t-value indicates that the observed mean is many standard
  errors away from the hypothesised mean, providing stronger evidence against
  the null hypothesis.

The t-statistic has an associated p-value, which quantifies how likely it is to
observe a t-value this extreme (or more extreme) if the null hypothesis were
true.

**Degrees of Freedom—(d.f.):** Degrees of freedom refers to the number of values
in a calculation that are free to vary. It is essentially a measure of how much
independent information is available for estimating a statistical parameter.

For example: Imagine you need to calculate the average height of 5 people, and
you know the sum of all their heights is 340 inches. If you know the heights of
4 of these people (65, 70, 68, and 72 inches), you can automatically determine
the height of the fifth person without measuring them: 340 - (65 + 70 + 68 + 72)
= 65 inches In this example, even though there are 5 people, you only have 4
degrees of freedom because once you know 4 heights and the total, the 5th height
is no longer “free to vary” – it is determined by the other values.

**Confidence Interval**

A 95% confidence interval (CI) for a parameter provides a range that, under
repeated sampling, would contain the true (but unknown) population parameter 95%
of the time. Analogy: Imagine shooting arrows at a target. If you drew a circle
around where 95% of your arrows landed, that circle is like a confidence
interval—it captures the region in which your “shots” (i.e., estimates from
different samples) tend to fall.

**Uncertainty quantification:** A CI communicates your estimate’s
precision—narrower intervals imply more precise estimates (often due to larger
samples or less variability), whereas wider intervals indicate greater
uncertainty about the true value.

**Academic Reporting (based on the APA 7th Edition Style)**

Below are some key considerations to note when reporting statistical analysis
using the APA style:

1.  The type of statistical test must be stated.

2.  Although not mandatory, the dependent variable is usually stated first
    followed by the independent variable when describing relationships, e.g.,
    “…to examine whether advertising expenditures on YouTube, TikTok, and
    Facebook collectively predict Sales” such that Sales is the dependent
    variable that depends on advertising expenditures on YouTube, TikTok, and
    Facebook.

3.  Test statistic and parameters: Report the appropriate test statistic
    (*t*-Statistic, *F*-Statistic, $\chi^2$ , etc.) with the degrees of freedom
    in parentheses. The italicized statistical symbol is immediately followed by
    the degrees of freedom in parentheses without a space, e.g., *t*(498) and
    not t (498).

4.  Exact p-values: Report exact p-values, when possible (e.g., *p* = .032),
    unless they are less than .001, then report as *p* \< .001.

5.  Effect sizes: Include appropriate effect size measures (e.g., R²) to
    indicate practical significance.

6.  Standard errors: Report standard errors of estimates when relevant. The
    standard error tells you how much your estimate might vary if you were to
    repeat your study with different random samples from the same population. A
    smaller standard error indicates a more precise estimate.

7.  Confidence Intervals (CI): The confidence level should be clearly stated
    whenever you report point estimates (e.g., means, regression coefficients,
    correlations, etc.). The 95% confidence interval is the most common, and if
    another level is used (e.g., 90% CI, 99% CI), it should be explicitly
    mentioned. Confidence intervals are typically enclosed in square brackets
    \[\], with the lower and upper limits separated by a comma. For example: 95%
    CI \[-.03, .04\]. They are usually reported directly after the statistic
    they describe, often within the same sentence or in parentheses.

8.  Two decimal places: Report to two decimal places, except p-values which may
    need three or more decimal places. Round to the nearest value; do not
    truncate.

9.  Descriptive statistics: Report relevant means, standard deviations, and
    sample sizes, e.g., The sample size included 500 observations (M = 25.43, SD
    = 4.62).

10. Italicize statistical symbols: Use italics for statistical symbols (*t*,
    *F*, *p*, etc.) but not for Greek letters ( $\mu$, $\sigma$, $\alpha$),
    subscripts, or parenthetical information, e.g., R² = .45, *F*(2, 97) =
    15.62, *p* \< .001, The participants (*N* = 120) had an average score (*M* =
    25.43, *SD* = 4.62) on the cognitive test.

Further reading: <https://apastyle.apa.org/jars>

## One-Way Analysis of Variance

### Technical Academic Report (Based on APA Guidelines of Reporting Results)

A one-way analysis of variance was conducted to compare the effect of the chosen
algorithm (Logistic Regression, Random Forest, Support Vector Machine, Gradient
Boosting) on classification accuracy. The effect of the chosen algorithm was
statistically significant, *F*(3, 196) = 104.70, *p* \< .001, η² = .62,
indicating that algorithm choice accounted for approximately 62% of the variance
in accuracy — a large effect by conventional benchmarks.

| *An omnibus test is a statistical test that evaluates whether there is a significant effect or difference somewhere among three or more groups or variables, without identifying which specific groups or pairs are driving that effect. The term "omnibus" (Latin for "for all") reflects that the test addresses a single overall question across the entire set of groups simultaneously, rather than each pairwise comparison individually.*

Prior to interpreting the omnibus test, assumptions were checked. Levene's test
indicated no violation of homogeneity of variance, *F*(3, 196) = 0.26, *p* =
.857. The Shapiro-Wilk test indicated that residuals did not significantly
depart from normality, *W* = .99, *p* = .099, though this value approaches the
conventional threshold and warrants visual confirmation given the sample size.
See the figure below for the Q-Q plot that provides a visual confirmation.

```{r Q-Q_plot_for_result}
#| echo: false
#| message: false
#| warning: false

plot(one_way_anova_result, which = 2)
```

A Durbin-Watson test found no evidence of positive autocorrelation in the
residuals, *DW* = 2.16, *p* = .829; this result is meaningful only insofar as
the row order of the dataset reflects a genuine sequence.

Post hoc comparisons using Tukey's HSD test were conducted to examine all
pairwise differences. Every comparison was statistically significant, producing
a complete and unambiguous ranking: Gradient Boosting \> Random Forest \>
Support Vector Machine \> Logistic Regression.

| Comparison | Mean Diff. | 95% CI | *p* (adj.) | Cohen's *d*¹ |
|----|----|----|----|----|
| Gradient Boosting vs. Logistic Regression | 0.090 | \[0.075, 0.105\] | \< .001 | 3.20 |
| Gradient Boosting vs. SVM | 0.069 | \[0.054, 0.083\] | \< .001 | 2.44 |
| Gradient Boosting vs. Random Forest | 0.027 | \[0.012, 0.041\] | \< .001 | 0.95 |
| Random Forest vs. Logistic Regression | 0.063 | \[0.049, 0.078\] | \< .001 | 2.25 |
| Random Forest vs. SVM | 0.042 | \[0.027, 0.057\] | \< .001 | 1.49 |
| SVM vs. Logistic Regression | 0.021 | \[0.007, 0.036\] | .001 | 0.76 |

¹*d* computed as mean difference divided by the pooled residual SD (√MSE =
.0281); not part of the original output, included for effect-size reporting.

### Non-Technical Business Report

Four predictive algorithms were tested head-to-head on the same accuracy
benchmark. The result is unambiguous: Gradient Boosting is the top performer,
Random Forest is a clear second, Support Vector Machine is third, and Logistic
Regression trails all three. The gap between the best and worst performer is
approximately nine percentage points of accuracy, and statistical testing
confirms this gap is real, not a product of random variation in the sample. All
underlying data-quality checks passed, so the result can be relied upon as
stated. The open question is not whether Gradient Boosting wins on accuracy — it
does, decisively — but whether accuracy alone is the right basis for a
deployment decision, given that cost, speed, and interpretability were not part
of this comparison.

### **Recommendations**

1.  **Shortlist Gradient Boosting** as the lead candidate for further
    development, on the strength of a statistically validated and practically
    meaningful accuracy advantage over all alternatives.

2.  **Do not finalize the deployment decision on accuracy alone.** Gather
    comparative data on training cost, inference latency, and interpretability
    before committing, since these factors were outside the scope of this
    analysis and may offset the accuracy gain depending on the application.

3.  **Confirm comparison fairness retrospectively**: verify that all four
    algorithms received equivalent tuning effort and identical data splits. An
    accuracy advantage attributable to unequal tuning budgets would not be a
    genuine algorithm effect.

4.  **Replicate on a held-out or temporally distinct dataset** before treating
    this ranking as generalizable, since a single benchmark run cannot rule out
    dataset-specific idiosyncrasy.

5.  **Document the data-generating process** behind the 200 accuracy
    observations (e.g., repeated k-fold cross-validation, repeated random
    splits) — this directly affects whether the independence assumption
    underlying the reported *p*-values holds.

## Two-Way Analysis of Variance

### Technical Academic Report (Based on APA Guidelines of Reporting Results)

A two-way analysis of variance was conducted to examine the effects of Algorithm
(Logistic Regression, Random Forest, Support Vector Machine, Gradient Boosting),
Dataset (10 distinct datasets), and their interaction on classification
accuracy.

Assumption checks supported the validity of the model. Levene's test indicated
homogeneity of variance across all 40 algorithm-by-dataset cells, *F*(39, 160) =
0.91, *p* = .629. The Shapiro-Wilk test indicated that residuals did not depart
from normality, *W* = 0.99, *p* = .587. A Durbin-Watson test found no evidence
of positive autocorrelation, *DW* = 2.11, *p* = .786.

| Source              | *df* | *SS*  | *MS* | *F*   | *p*     | η²  |
|---------------------|------|-------|------|-------|---------|-----|
| Algorithm           | 3    | 0.25  | 0.08 | 99.18 | \< .001 | .62 |
| Dataset             | 9    | 0.01  | 0.00 | 0.81  | .612    | .02 |
| Algorithm × Dataset | 27   | 0.02  | 0.00 | 0.68  | .881    | .04 |
| Residuals           | 160  | 0.133 | 0.00 | —     | —       | .33 |

The main effect of Algorithm was statistically significant and large, *F*(3,
160) = 99.18, *p* \< .001, η² = .616. The main effect of Dataset was not
significant, *F*(9, 160) = 0.81, *p* = .612, η² = .02, indicating that accuracy
did not differ systematically across the ten datasets when averaged over
algorithm. The Algorithm × Dataset interaction was not significant, *F*(27, 160)
= 0.68, *p* = .881, η² = .04, indicating that the relative ordering of the four
algorithms did not vary meaningfully from one dataset to another.

**Post hoc comparisons: Algorithm.** Tukey's HSD confirmed that all six pairwise
algorithm differences were statistically significant, reproducing the strict
ordering Gradient Boosting \> Random Forest \> Support Vector Machine \>
Logistic Regression.

| Comparison                              | Diff. | 95% CI         | *p* (adj.) | *d*¹ |
|-----------------------------------------|-------|----------------|------------|------|
| Gradient Boosting – Logistic Regression | 0.09  | \[0.08, 0.11\] | \< .001    | 3.12 |
| Gradient Boosting – SVM                 | 0.07  | \[0.05, 0.08\] | \< .001    | 2.38 |
| Gradient Boosting – Random Forest       | 0.03  | \[0.01, 0.04\] | \< .001    | 0.93 |
| Random Forest – Logistic Regression     | 0.06  | \[0.05, 0.08\] | \< .001    | 2.20 |
| Random Forest – SVM                     | 0.04  | \[0.03, 0.06\] | \< .001    | 1.46 |
| SVM – Logistic Regression               | 0.02  | \[0.01, 0.04\] | .002       | 0.74 |

¹*d* = mean difference / √MSE (√0.00083 = 0.0288); included for effect-size
context, not part of original output.

**Post hoc comparisons: Dataset.** Consistent with the non-significant main
effect, none of the 45 pairwise dataset comparisons reached significance (*p*
adj. range = .838–1.000).

**Post hoc comparisons: Algorithm × Dataset.** Given the non-significant omnibus
interaction, the 780 individual cell-level contrasts are not interpreted or
reported in full. A number of these contrasts nonetheless show small adjusted
*p*-values (e.g., Gradient Boosting:Dataset_1 vs. Logistic Regression:Dataset_2,
*p* = .0002); this pattern reflects the large main effect of Algorithm surfacing
within individual cell pairs, not a true interaction, and should not be read as
evidence that algorithm performance depends on dataset. Standard practice is to
withhold interpretation of simple effects when the omnibus interaction test is
non-significant, and that practice is followed here.

### Non-Technical Business Report

The earlier finding — that Gradient Boosting outperforms the other three
algorithms — has now been stress-tested across ten distinct datasets, and it
holds without exception. Performance does not vary meaningfully by dataset on
its own, and, critically, the ranking of algorithms (Gradient Boosting first,
then Random Forest, then SVM, then Logistic Regression) is identical across all
ten datasets — there is no scenario in this evidence base where a different
algorithm comes out on top. This substantially strengthens the earlier
conclusion: the advantage is not a quirk of one dataset or one lucky sample, but
a consistent pattern. All data-quality checks passed. The remaining decision is
the same one flagged previously: accuracy is settled, but cost, latency, and
interpretability still need to be considered before committing to a production
choice.

### Recommendations

1.  **Proceed with Gradient Boosting as the lead candidate** with higher
    confidence than before — the result is now supported by consistency across
    ten datasets, not a single benchmark run.

2.  **Establish precisely what the ten "datasets" represent** before treating
    this as evidence of cross-domain generalization. This single clarification
    materially changes how strong a claim can be made.

3.  **Still price in cost, latency, and interpretability** before finalizing
    deployment, as these remain untested dimensions.

4.  **Do not interpret the individual Algorithm × Dataset cell comparisons** in
    any downstream reporting; only the omnibus interaction test
    (non-significant) is the supportable claim.

5.  **Validate on at least one genuinely held-out dataset** collected
    independently of the original ten, to confirm the result is not an artifact
    of how those ten datasets were originally selected or constructed.

## Multiple Analysis of Variance

### Technical Academic Report (Based on APA Guidelines of Reporting Results)

A one-way multivariate analysis of variance (MANOVA) was conducted to examine
the effect of Algorithm (Logistic Regression, Random Forest, Support Vector
Machine, Gradient Boosting) on a combined outcome of classification Accuracy,
Precision, and Recall.

**Assumption checks.** Mardia's test indicated no significant departure from
multivariate normality in skewness, *b* = 10.71, *p* = .381, or kurtosis, *b* =
0.28, *p* = .782. Box's M test indicated no significant violation of the
homogeneity of covariance matrices assumption, χ²(18) = 12.36, *p* = .828. A
Durbin-Watson test found no evidence of positive autocorrelation, *DW* = 2.16,
*p* = .815.

**Multivariate effect.** The multivariate effect of Algorithm on the combined
outcome was statistically significant, Pillai's Trace = 0.67, *F*(9, 588) =
18.60, *p* \< .001.

**Univariate follow-up tests.** Given the significant multivariate effect,
univariate ANOVAs were examined for each outcome as protected follow-up tests.

| Outcome   | *df*   | *MS*   | *F*    | *p*     | η²   |
|-----------|--------|--------|--------|---------|------|
| Accuracy  | 3, 196 | 0.0823 | 104.75 | \< .001 | .616 |
| Precision | 3, 196 | 0.0833 | 74.55  | \< .001 | .533 |
| Recall    | 3, 196 | 0.0838 | 77.71  | \< .001 | .543 |

All three effects were significant and large, with Algorithm explaining between
53% and 62% of variance depending on outcome.

**Post hoc comparisons.** Tukey's HSD was conducted separately for each outcome.

*Accuracy*

| Comparison                              | Diff. | 95% CI         | *p* (adj.) | *d*¹ |
|-----------------------------------------|-------|----------------|------------|------|
| Gradient Boosting – Logistic Regression | 0.09  | \[0.08, 0.11\] | \< .001    | 3.21 |
| Gradient Boosting – SVM                 | 0.07  | \[0.05, 0.08\] | \< .001    | 2.45 |
| Gradient Boosting – Random Forest       | 0.03  | \[0.01, 0.04\] | \< .001    | 0.95 |
| Random Forest – Logistic Regression     | 0.06  | \[0.05, 0.08\] | \< .001    | 2.26 |
| Random Forest – SVM                     | 0.04  | \[0.03, 0.06\] | \< .001    | 1.50 |
| SVM – Logistic Regression               | 0.02  | \[0.01, 0.04\] | .001       | 0.76 |

*Precision*

| Comparison                              | Diff. | 95% CI         | *p* (adj.) | *d*¹ |
|-----------------------------------------|-------|----------------|------------|------|
| Gradient Boosting – Logistic Regression | 0.09  | \[0.07, 0.11\] | \< .001    | 2.64 |
| Gradient Boosting – SVM                 | 0.07  | \[0.05, 0.08\] | \< .001    | 2.00 |
| Gradient Boosting – Random Forest       | 0.02  | \[0.00, 0.04\] | .018       | 0.59 |
| Random Forest – Logistic Regression     | 0.07  | \[0.05, 0.09\] | \< .001    | 2.05 |
| Random Forest – SVM                     | 0.05  | \[0.03, 0.06\] | \< .001    | 1.40 |
| SVM – Logistic Regression               | 0.02  | \[0.00, 0.04\] | .008       | 0.65 |

*Recall*

| Comparison                              | Diff. | 95% CI         | *p* (adj.) | *d*¹ |
|-----------------------------------------|-------|----------------|------------|------|
| Gradient Boosting – Logistic Regression | 0.09  | \[0.07, 0.11\] | \< .001    | 2.69 |
| Gradient Boosting – SVM                 | 0.07  | \[0.05, 0.08\] | \< .001    | 1.99 |
| Gradient Boosting – Random Forest       | 0.02  | \[0.00, 0.04\] | .030       | 0.56 |
| Random Forest – Logistic Regression     | 0.07  | \[0.05, 0.09\] | \< .001    | 2.14 |
| Random Forest – SVM                     | 0.05  | \[0.03, 0.06\] | \< .001    | 1.43 |
| SVM – Logistic Regression               | 0.02  | \[0.01, 0.04\] | .003       | 0.71 |

¹*d* = mean difference / √MSE; not part of the original output, included for
effect-size context.

**Synthesis.** Across all three outcomes, every pairwise comparison was
statistically significant, and the rank ordering was identical: Gradient
Boosting \> Random Forest \> SVM \> Logistic Regression. This convergence across
three distinct, though related, performance metrics provides stronger evidence
for the algorithm ranking than any single outcome could provide alone. The
Gradient Boosting–Random Forest contrast was consistently the smallest margin
among the six comparisons in every outcome and is identified below as the
comparison most sensitive to multiplicity correction.

### Non-Technical Business Report

The conclusion is now fully confirmed: Gradient Boosting leads, Random Forest is
second, SVM third, Logistic Regression last — and this exact ordering holds
independently across all three quality measures (accuracy, precision, and
recall), with every single pairwise difference statistically significant in
every measure. This is the most complete evidence produced across this
evaluation: no algorithm wins on one metric while losing on another, and no
result depends on which single metric is chosen to judge the algorithms. The
only nuance worth flagging at the leadership level is that the gap between the
top two performers — Gradient Boosting and Random Forest — is consistently the
narrowest of the six comparisons across all three metrics. The overall ranking
is not in question; how decisively Gradient Boosting separates itself from the
runner-up is the one place where the margin is closer than the headline result
suggests.

### Recommendations

1.  **Proceed with Gradient Boosting as the recommended algorithm.** The
    conclusion is now supported by full convergence across three independent
    performance metrics, the strongest evidentiary standard reached in this
    evaluation.

2.  **Treat the Gradient Boosting vs. Random Forest margin as the genuine
    decision point**, not the broader ranking. If training cost, inference
    latency, or interpretability favor Random Forest, this is the one comparison
    where that trade-off is close enough to matter.

3.  **Apply a stricter correction across the three outcome families** (e.g.,
    Bonferroni at α = .0167 instead of .05) specifically to the Gradient
    Boosting–Random Forest contrasts before treating them as fully secure, since
    Tukey's correction here only protects within each individual metric's six
    comparisons, not across all eighteen.

4.  **Report all three Tukey tables together**, as done above, rather than any
    one in isolation — the convergence across metrics is itself part of the
    evidentiary strength and should not be dropped in summarized reporting.

5.  **If forced to choose between Gradient Boosting and Random Forest under
    resource constraints**, treat it as a genuine cost-benefit decision rather
    than a statistical formality, given how narrow and comparatively fragile
    that specific margin is.

# References and Further Reading

American Psychological Association. (2025, February). *Journal Article Reporting
Standards (JARS)*. APA Style. Retrieved April 28, 2025, from
<https://apastyle.apa.org/jars>

Hodeghatta, U. R., & Nayak, U. (2023). *Practical Business Analytics Using R and
Python: Solve Business Problems Using a Data-driven Approach* (2nd ed.). Apress.
<https://link.springer.com/book/10.1007/978-1-4842-8754-5>
